---
title: "``ols_fpe``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(olsrr)
```

### Final prediction error

#### Description 
 
Estimated mean square error of prediction.

####Usage 
<pre><code>

ols_fpe(model)

</code></pre>
#### Arguments 
 
* ``model``:  An object of class lm.



####Examples 
```{r} 
model <- lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_fpe(model)
```

#### Details 
 
Computes the estimated mean square error of prediction for each model selected assuming that the values of the regressors are fixed and that the model is correct.
MSE((n + p)=n)
where MSE = SSE=(n 􀀀 p), n is the sample size and p is the number of predictors including the intercept

#### Value
Final prediction error of the model.


#### See Also
Other model selection criteria: ols_aic, ols_apc, ols_hsp, ols_mallows_cp, ols_msep, ols_sbc,ols_sbic

#### References
Akaike, H. (1969). “Fitting Autoregressive Models for Prediction.” Annals of the Institute of Statistical
Mathematics 21:243–247.
Judge, G. G., Griffiths, W. E., Hill, R. C., and Lee, T.-C. (1980). The Theory and Practice of
Econometrics. New York: John Wiley & Sons.

