---
title: "Probability Distributions with R"
subtitle: "Introduction to R"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

### Probability Distribution 
A computer software company which specializes in statistical software sells 4 software licences every day, on average.

#### Exercises 
Answer the following questions.

i What is the probability that the software company sells at least two licence in one particular day?
ii What is the probability that the software company will sell exactly one licence in one particular day?
iii What is the probability that the software company will sell sixteen licences or more in a five day working week?

### Exercise 1


```{r}

dpois(0:4,lambda=2)
```

---

### Exercise 2


```{r}

dpois(0:4,lambda=2)
```

---

### Exercise 3


```{r}
dpois(15,lambda=10)

```