---
title: "Probability Distributions with R"
subtitle: "Introduction to R"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

Three per cent of PCB boards purchased over the Internet have some defect.

Consignments are sent in nominal batches of 100 PCB boards, but to account for the inclusion of defective items, there are 105 per consignment.

A consignment is opened and each PCB board is tested.
What is the probability that:

#### Exercises
1.  6 or more boards have some defect?
2.  Exactly 4 boards are defective?
3.  Exactly 6 boards are defective?
4.  Between 6 boards and 8 boards are defective?
4.  Between 2 boards and 4 boards are defective?
6.  If more than 8 boards from the batch were defective what action would you take, if any? (justify)
 
---

Binomial Probability Distribution

* $X$ is the number of defective components in the batch.
* $n$ is the number of independent trials. Here $n=105$.
* $p$ is the probability of a success. Here a success is the event that a selected component is found to be defective. $p=0.03$


$$X \sim \mbox{Bin}(n=105,p=0.03)$$

---

 
### Exercise 1

$P(X\geq 6) = 1 - P(X\leq 5)$

```{r}
n <- 105
p <- 0.03

1 -pbinom(5,size=n,prob=p)
```

---

### Exercise 2

$P(X=4)$

```{r}
dbinom(4,size=n,prob=p)

```

---

### Exercise 3

$P( X = 6)$

```{r}
dbinom(6,size=n,prob=p) 

```


---
### Exercise 4

$P( 6\leq X \leq 8$)

```{r}
pbinom(8,size=n,prob=p) - pbinom(5,size=n,prob=p)

```

---
### Exercise 5

$P( 2\leq X \leq 4$)

```{r}
pbinom(4,size=n,prob=p) - pbinom(1,size=n,prob=p)

```

---

### Exercise 6

$P(X > 8)$


```{r}
pbinom(8,size=n,prob=p,lower.tail = FALSE)

```

---


---