---
title: "Probability Distributions with R"
subtitle: "Introduction to R"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


```{r echo=FALSE}
# - HC Module 3 2012
# - RSS 2012
# - EXAMINATIONS OF THE HONG KONG STATISTICAL SOCIETY
# - HIGHER CERTIFICATE IN STATISTICS, 2012
# - MODULE 3 : Basic statistical methods
# - Question 2
```


2. (a) A researcher wishes to conduct an experiment to see whether or not consuming 
alcohol affects the mean reaction time of drivers.
(i) What would be a suitable null hypothesis, and an appropriate alternative 
hypothesis?

(ii) What would be the result of making a Type I error and what would be 
the result of making a Type II error?

(iii) If one researcher uses a significance level of = 0.05 and another uses 
a significance level of = 0.01, which would be more likely to make a 
Type I error and which would be more likely to make a Type II error?

---

A test of mental ability has been constructed so that, for adults in the UK, the 
test score is Normally distributed with mean 100 and standard deviation 15.
A doctor needs to test at the 5% significance level whether the mean score of 
sufferers from a particular disease differs from the mean score of the general 
population on this test. She has obtained the following test scores for ten 
randomly selected patients suffering from the disease.

$$119 131 95 107 125 90 128 89 103 103$$

---
#### Exercises 
(i) Perform the appropriate test and report your conclusions to the doctor.
(8)
(ii) Suppose that, before collecting data, the doctor had suspected that the 
disease led to a reduction in mental ability. How would that have 
affected your test and your conclusions?

(iii) What is the minimum sample size for which a sample mean test score 
of 109 would be indicative of a significant increase in mental ability at 
the 1% significance level?

---