---
title: "Statistal Inference"
subtitle: "RSS Module 3 Exam Paper 2011"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


```{r echo=FALSE}
# - HC Module 3 2011
# - RSS 2011
# - EXAMINATIONS OF THE HONG KONG STATISTICAL SOCIETY
# - HIGHER CERTIFICATE IN STATISTICS, 2011
# - MODULE 3 : Basic statistical methods
# - Question 1
```

### Fertilizers

Ten plants of a certain type, grown under standard conditions and treated with a new 
brand of fertiliser, attain the following heights (in cm).
<pre><code>
25 28 24 23 27 30 24 21 28 30
</code></pre>

* The sample mean is 26.0 and the sample variance is 9.33. 

* From extensive previous experiments, it is known that plants of the same type, grown under similar conditions but treated with a standard fertiliser, attain a mean height of 25.0 cm.


---

### Inputing Data into R environment

```{r}
Heights <- c(25,28,24,23,27,30,24,21,28,30)
```

```{r}
mean(Heights)
```

```{r}
var(Heights)
```


```{r}
sd(Heights)
```

---

### Exercises

***Questions on the RSS Exam Paper***

(i) Write down a statistical model for the distribution of plant heights in the 
population from which the sample is assumed to be drawn, including the 
definition of any unknown parameters.

(ii) Obtain a 99% confidence interval for the population mean height of plants 
treated with the new fertiliser.

(iii) Carefully specifying your hypotheses, conduct a test at the 10% significance 
level to decide whether there is evidence to suggest that the new fertiliser 
produces plants of greater mean height than the standard fertiliser. State your 
conclusions.

(iv) Provide a 95% confidence interval for the population variance of the heights of 
plants treated with the new brand of fertiliser.

---

### 99% Confidence Interval

We will use $\mu=25$ here

```{r}
t.test(Heights,mu=25,alpha=0.01)
```
The Confidence interval is (23.8,28.2)

---

### Using <tt> tidy() </tt>

The <tt>tidy()</tt> function from the {broom} R package

```{r}
library(broom)
t.test(Heights,mu=25,alpha=0.01) %>% tidy()
```



---


### Hypothesis Test


```{r}
t.test(Heights,mu=25,
       alpha=0.10,
       alternative="two.sided")
```

---

```{r}
t.test(Heights,mu=25,
       alpha=0.10,
       alternative="greater")
```

---

```{r}
library(broom)
t.test(Heights,mu=25,alpha=0.10,alterative="greater") %>% tidy()
```

---

### <tt>Desctools::VarCI()</tt>

#### Confidence Intervals for the Variance

Calculates confidence intervals for the variance. 

Available approachs are the classical one using the ChiSquare distribution, a more robust version proposed by Bonett and the bootstrap options available in the package boot.

#### Usage

<pre><code>VarCI(x, 
      method = c("classic", "bonett", "norm", 
                  "basic", "stud", "perc", "bca"),
      conf.level = 0.95, 
      sides = c("two.sided", "left", "right"),
      na.rm = FALSE, R = 999)
</code></pre>      

---

#### 95% Confidence Interval for Variance using standard method
```{r warning=FALSE,message=FALSE}
library(DescTools)
VarCI(Heights)
```

#### 99% Confidence Interval for Variance using "bonett" method
```{r warning=FALSE,message=FALSE}
library(DescTools)
VarCI(Heights,
      method="bonett",
      conf.level = 0.95)
```

---

---