
---
title: "Probability Distributions with R"
subtitle: "Introduction to R"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


```{r echo=FALSE}
# - HC Module 3 2011
# - RSS 2011
# - EXAMINATIONS OF THE HONG KONG STATISTICAL SOCIETY
# - HIGHER CERTIFICATE IN STATISTICS, 2011
# - MODULE 3 : Basic statistical methods
# - Question 2
```

2. Twenty cows were used in an experiment to compare two types of feed, Feed A and 
Feed B. Half of the cows, chosen at random, were fed with Feed A over a certain 
period of time and the other half with Feed B. In fact two of the cows on Feed B were 
wrongly fed for part of the period and they had to be removed from the experiment. 
The gains in weight, in pounds, of the remaining cows over the period of the 
experiment are listed below. The corresponding sample means and variances are also 
given.
Feed A 30 26 30 19 25 37 27 38 26 31
Feed B 40 34 28 29 26 36 28 37
sample mean sample variance
Feed A 28.9 32.1
Feed B 32.3 26.5
(i) Defining any unknown parameters, write down a statistical model for the 
distributions of weight gains in the two populations of cows given the two 
feeds, assuming that the variances of these two distributions are equal.
(3)
(ii) Carefully specifying your hypotheses, test at the 5% significance level whether 
there is a difference between the mean weight increases for the two feeds.
(10)
(iii) It is now required to test at the 5% significance level the assumption that the 
variances of the two weight gain distributions are equal. State any adjustments 
that are needed to the model in part (i), carefully specify the hypotheses to be 
tested, carry out the test and give your conclusions.
