---
title: "Probability Distributions with R"
subtitle: "Introduction to R"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

Assume that the number of weekly study hours for students at a certain university is approximately normally distributed with a mean of 22 and a standard deviation of 6.

#### Exercises

1. Find the probability that a randomly chosen student studies less than 12 hours.

2. Estimate the percentage of students that study more than 37 hours.

---

### Solution
$X$ is number of weekly study hours, a normally distributed random variable with $\mu=22$ and standard deviation $\sigma = 6$

$$X \sim N(\mu =22,\sigma^2 =36) $$


```{r}

mu <- 22
sigma <-  6
```


---

## {DescTools R package}


**DescTools - Tools for Descriptive Statistics**

 
A collection of miscellaneous basic statistic functions and convenience wrappers for efficiently describing data. The author's intention was to create a toolbox, which facilitates the (notoriously time consuming) first descriptive tasks in data analysis, consisting of calculating descriptive statistics, drawing graphical summaries and reporting the results.

***Maintainer*** - Andri Signorelli

```{r warning = FALSE}

library(DescTools)

```

---

#### Compute the Z-Score

$$Z = \frac{ X - \mu}{\sigma}$$


* $P(X\leq x_{o}) = P(Z \leq z_{o})$
* $P(Z \leq -z_{o} = P(Z \geq z_{o})$

---

### Exercise 1

Find the probability that a randomly chosen student studies less than 12 hours.

```{r}

pnorm(12,mean=mu,sd=sigma)
```

### Z-score

```{r}
(12-mu)/sigma
```
---

using **Shade()** function from **{DescTools}** R package

<pre><code>
curve(dnorm(x, mean=22, sd=6), xlim=c(2,42),
      main=paste("Normal Distribution Probability Density Function"),
      type="n", las=1, ylab="probability", xlab="Study Hours (X)")

Shade(dnorm(x, mean=22, sd=6), breaks=c(2,12,42),
      col=c(hred, hblue), density=c(35, 10))
</code></pre>



---

```{r echo=FALSE}
curve(dnorm(x, mean=22, sd=6), xlim=c(2,42),
      main=paste("Normal Distribution Probability Density Function"),
      type="n", las=1, ylab="probability", xlab="Study Hours (X)")

Shade(dnorm(x, mean=22, sd=6), breaks=c(2,12,42),
      col=c(hred, hblue), density=c(35, 10))
```

---


### Exercise 2

Estimate the percentage of students that study more than 37 hours.

```{r}

1- pnorm(37,mean=mu,sd=sigma)
```

```{r}

pnorm(37,mean=mu,sd=sigma,lower.tail = FALSE)
```

---

#### Z-Score

```{r}
(37-mu)/sigma

```
---

<pre><code>
curve(dnorm(x, mean=22, sd=6), xlim=c(2,42),
      main=paste("Normal Distribution Probability Density Function"),
      type="n", las=1, ylab="probability", xlab="Study Hours (X)")

Shade(dnorm(x, mean=22, sd=6), breaks=c(2,37,42),
      col=c( hblue,hred), density=c(10,35))
</code></pre>

---

```{r echo=FALSE}
curve(dnorm(x, mean=22, sd=6), xlim=c(2,42),
      main=paste("Normal Distribution Probability Density Function"),
      type="n", las=1, ylab="probability", xlab="Study Hours (X)")

Shade(dnorm(x, mean=22, sd=6), breaks=c(2,37,42),
      col=c( hblue,hred), density=c(10,35))
```

---

---
