---
title: "Statistics with R"
subtitle: "Probability Distributions"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


The life-time (in days) of a certain electronic component that operates in a high-temperature environment is log-normally distributed with $\mu=1.2$ and $\sigma=0.5$.

Using a seed of 1234, simulate the lifetimes of 10,000 such components. Based on this sample, answer the following questions. 

#### Questions 

a. Estimate the mean and variance of the lifetimes of the electronic component.

b. Estimate the probability that the component works no more than 4 days.

c. Estimate the probability that the component works more than 5 days.

d. Estimate the probability that the component works between 3 and 5 days.

---

## The Log Normal Distribution

####Description

Density, distribution function, quantile function and random generation for the log normal distribution whose logarithm has mean equal to meanlog and standard deviation equal to sdlog.

#### Usage
<pre><code>
dlnorm(x, meanlog = 0, sdlog = 1, log = FALSE)
plnorm(q, meanlog = 0, sdlog = 1, lower.tail = TRUE, log.p = FALSE)
qlnorm(p, meanlog = 0, sdlog = 1, lower.tail = TRUE, log.p = FALSE)
rlnorm(n, meanlog = 0, sdlog = 1)
</code></pre>

---

### R environment and packages

```{r include= FALSE,echo=FALSE,warning=FALSE}

library(tidyverse)
library(magrittr)
options(digits=4)
```


<pre><code>

library(tidyverse)
library(magrittr)

options(digits=4)
</code></pre>


```{r}
set.seed(1234)
```
---


### Simulation of 10,000 component lifetimes

```{r}


#Number of Simulations
N <- 10000

Lifetimes <- rlnorm(n = N , meanlog = 1.2, sdlog = 0.5)

head(Lifetimes)
```

---

### Histogram of the simulated lifetime values

```{r}

hist(Lifetimes,breaks=200)

```


---

### Part (a)

#### Estimates based on the simulated data

```{r}

mean(Lifetimes)

var(Lifetimes)

sd(Lifetimes)


```

---

### Population Mean and Variance

\begin{eqnarray*}
E(X)&=&e^{\mu+\sigma^2/2}\\
& & \\
&=&e^{1.2+(0.5^2)/2} = e^{1.325}\\
& & \\
&=&3.7622\\
& & \\
& & \\
\operatorname{Var}(X)&=&e^{2\mu+\sigma^2}(e^{\sigma^2−1})\\& & \\
&=&e^{(2\times 1.2)+0.5^2}\cdot(e^{0.5^2−1})\\& & \\
&=&e^{2.65}\cdot (e^{0.25−1}) = 14.154\cdot (0.284)\\& & \\
&=&4.0197\\& & \\
\end{eqnarray*}
---


```{r}
median(Lifetimes)
```


```{r}
qlnorm(0.5,meanlog = 1.2,sdlog = 0.5)
```




---

### Part (b)

Estimate the probability that the component works no more than 4 days, $P(X \leq 4)$

#### Using the simulated data

```{r}

mean(Lifetimes <4)

```



#### Using the cumulative distribution function

```{r}

plnorm(4 , meanlog = 1.2, sdlog = 0.5)
```

---


### Part (c)

Estimate the probability that the component works more than 5 days, $P(X \geq 5)$

#### Using the simulated data

```{r}

mean(Lifetimes >5)

```

#### Using the cumulative distribution function

```{r}

plnorm(5 , meanlog = 1.2, sdlog = 0.5, lower.tail = FALSE)
```


---


### Part (d)

Estimate the probability that the component works between 3 and 5 days, $P(3 \leq X \leq 5)$


```{r}
# using dplyr::between

mean( between(Lifetimes, 3, 5) )
        
        
```        

#### Using the cumulative distribution function

```{r}

plnorm(5 , meanlog = 1.2, sdlog = 0.5) - 
  plnorm(3 , meanlog = 1.2, sdlog = 0.5)
```

