---
title: "Probability Distributions with R"
subtitle: "Introduction to R"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(DescTools)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

A machine is said to have a mean failure free lifetime of 2000 hours. It is assumed that the failure free lifetimes of the machines have an exponential distribution. What is the probability that the machine will last:

#### Exercises 

1. at least 1000 hours
2. at least 1500 hours
3. less than 800 hours
4. between 1300 hours and 1800 hours

---

#### Exponential Distribution 

An exponentially distributed random variable $X$ is specifed with the ***rate parameter*** $\lambda$
$$ X \sim \mbox{exp}(\lambda)  $$
The  Expected Value of X is computed using the following formula
$$ E(X) = \frac{1}{\lambda} $$

---

#### Probability Density Function 

 The probability density function (PDF) of an exponential distribution is

$${ \displaystyle f(x;\lambda) = 
\lambda e^{-\lambda x} \mbox{ where }x \geq 0 }$$

 The parameter $\lambda$  is called ***rate parameter***. It is the inverse of the expected duration ($\mu$).

 (If the expected duration is 5 (e.g. five minutes) then the rate parameter value is 0.2.)



---

## {DescTools R package}


**DescTools - Tools for Descriptive Statistics**

 
A collection of miscellaneous basic statistic functions and convenience wrappers for efficiently describing data. The author's intention was to create a toolbox, which facilitates the (notoriously time consuming) first descriptive tasks in data analysis, consisting of calculating descriptive statistics, drawing graphical summaries and reporting the results.

***Maintainer*** - Andri Signorelli

```{r warning = FALSE}

library(DescTools)

```
---

### Exercise 1

Probability of the failure-free lifetime is at least 1000 hours.i.e. $P(X\geq 1000)$

```{r}
lambda <- 2000
```

#### Cumulative Distribution Function

```{r}
pexp(1000, rate =(1/lambda) )

```

$P(X\geq 1000)$

```{r}
1- pexp(1000, rate =(1/lambda) )

```


---

<pre><code>
curve(dexp(x, rate =(1/lambda) ), 
      xlim=c(0,6000),
      main=paste("Exponent Distribution PDF"),
      type="n", las=1, 
      ylab="probability", 
      xlab="Failure Free Lifetime (X)")

Shade(dexp(x, rate =(1/lambda) ), 
      breaks=c(0,1000,6000),
      col=c( hblue, hred ), 
      density=c(10,45))
</code></pre>

---

```{r echo=FALSE}
curve(dexp(x, rate =(1/lambda) ), xlim=c(0,6000),
      main=paste("Exponent Distribution PDF"),
      type="n", las=1, ylab="probability", xlab="Failure Free Lifetime (X)")

Shade(dexp(x, rate =(1/lambda) ), breaks=c(0,1000,6000),
      col=c( hblue, hred ), density=c(10,45))
```

---

### Exercise 2

Probability of the failure-free lifetime is at least 1500 hours. 
$P(X\geq 1500)$

```{r}

pexp(1500,rate=(1/lambda),lower.tail = FALSE)
```

---

<pre><code>
curve(dexp(x, rate =(1/lambda) ), 
      xlim=c(0,6000),
      main=paste("Exponent Distribution PDF"),
      type="n", las=1, 
      ylab="probability", 
      xlab="Failure Free Lifetime (X)")

Shade(dexp(x, rate =(1/lambda) ), 
      breaks=c(0,1500,6000),
      col=c( hblue, hred ), 
      density=c(10,45))
</code></pre>

---

```{r echo=FALSE}
curve(dexp(x, rate =(1/lambda) ), xlim=c(0,6000),
      main=paste("Exponent Distribution PDF"),
      type="n", las=1, ylab="probability", xlab="Failure Free Lifetime (X)")

Shade(dexp(x, rate =(1/lambda) ), breaks=c(0,1500,6000),
      col=c( hblue, hred ), density=c(10,45))
```



---

### Exercise 3

Probability that the Failure Free Lifetime is less than 800 hours

$$P(X \leq 800)$$

```{r}

pexp(800, rate= (1/lambda))
```

---

<pre><code>
curve(dexp(x, rate =(1/lambda) ), 
      xlim=c(0,6000),
      main=paste("Exponent Distribution PDF"),
      type="n", las=1, 
      ylab="probability", 
      xlab="Failure Free Lifetime (X)")

Shade(dexp(x, rate =(1/lambda) ), 
      breaks=c(0,900,6000),
      col=c( hred, hblue), 
      density=c(45,10))
</code></pre>

---

```{r echo=FALSE}
curve(dexp(x, rate =(1/lambda) ), xlim=c(0,6000),
      main=paste("Exponent Distribution PDF"),
      type="n", las=1, ylab="probability", xlab="Failure Free Lifetime (X)")

Shade(dexp(x, rate =(1/lambda) ), breaks=c(0,900,6000),
      col=c( hred, hblue), density=c(45,10))
```

---


#### Exercise 4

Probability that the Failure Free Lifetime is between 1300 hours and 1800 hours $P(1300\leq X \leq 1800)$

```{r}

pexp(1800,rate=(1/lambda)) - pexp(1300,rate=(1/lambda))

```

---

<pre><code>
curve(dexp(x, rate =(1/lambda) ), 
      xlim=c(0,6000),
      main=paste("Exponent Distribution PDF"),
      type="n", las=1, 
      ylab="probability", 
      xlab="Failure Free Lifetime (X)")

Shade(dexp(x, rate =(1/lambda) ), 
      breaks=c(0,1300,1800,6000),
      col=c(hblue, hred, hblue), 
      density=c(10,45,10))
</code></pre>

---

```{r echo=FALSE}
curve(dexp(x, rate =(1/lambda) ), xlim=c(0,6000),
      main=paste("Exponent Distribution PDF"),
      type="n", las=1, ylab="probability", xlab="Failure Free Lifetime (X)")

Shade(dexp(x, rate =(1/lambda) ), breaks=c(0,1300,1800,6000),
      col=c(hblue, hred, hblue), density=c(10,45,10))
```

---


---


---

%------------------------------------------------------------------------%

\subsection*{Cumulative Distribution Function}
 The cumulative distribution function (CDF) of an exponential distribution is

\
P(X \leq x) = F(x) = \begin{cases}
1-e^{-\lambda x}, & x \ge 0, 
0, & x < 0.
\end{cases}\

 (\textbf{Important}) The CDF can be written as the probability of the lifetime being less than some value $x$.

\ P(X \leq x) = 1-e^{-\lambda x} \



 The complement of the CDF (i.e. $P(X \geq x)$ is

\
P(X \geq x) = \begin{cases}
e^{-\lambda x}, & x \ge 0, 
0, & x < 0.
\end{cases}\

\newpage 







%------------------------------------------------------------------------%

\subsection*{Expected Value and Variance}
%---------------------------------------------------------------------%

Here $\lambda > 0$ is the parameter of the distribution, often called the \textbf{rate parameter}. 

The distribution is supported on the interval $0, \infty)$.The expected value $E(X)$ of an exponentially distributed random variable $X$, specifed with the \textbf{rate parameter} $\lambda$
\ X \sim \mbox{exp}(\lambda)  \
is computed using the following formula
\ E(X) = \frac{1}{\lambda}. \
The expected value is also known as the exponential mean $\mu$.
\smallskip
 The variance of an exponential random variable $X$ is:
\\operatorname{Var}(X) = \frac{1}{\lambda^2}\

 Suppose $\lambda=0.1$


\
E(X) = \frac{1}{\lambda} = \frac{1}{0.1} = 10 \
The variance of an exponential random variable $X$ is:

\\operatorname{Var}(X) = \frac{1}{\lambda^2} = 100\

