---
title: "Probability Distributions with R"
subtitle: "Introduction to R"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

```{r echo=FALSE}
# - HC Module 3 2012
# - RSS 2012
# - EXAMINATIONS OF THE HONG KONG STATISTICAL SOCIETY
# - HIGHER CERTIFICATE IN STATISTICS, 2012
# - MODULE 3 : Basic statistical methods
# - Question 4
```

4. It is of interest to examine whether Choice Reaction Time (CRT) of table-tennis 
players is reduced by training.
(a) Sixteen table-tennis players took part in a trial in which each was randomly 
assigned to one of two groups. The first group of eight had not received 
training, and the second group of eight had received training. The CRT values 
below are each in hundredths of a second.
Group 1 (no training): 82, 69, 73, 44, 58, 56, 76, 66
Group 2 (after training): 63, 42, 74, 37, 50, 43, 80, 64

---
#### Exercises 

(i) Explain why it may be deemed appropriate to use the Wilcoxon rank 
sum test, and state what null and alternative hypotheses would be 
tested.
(3)
(ii) Carry out the Wilcoxon rank sum test at the 5% level of significance, 
and report your conclusions.
(6)
(b) Suppose instead that CRT was measured before and after training on eight 
randomly selected table-tennis players, with the results given below.
Player x (before training) y (after training)
1 82 63
2 69 42
3 73 74
4 44 37
5 58 50
6 56 43
7 76 80
8 66 64
(i) Explain why it may be deemed appropriate to use the Wilcoxon signedrank test, and state what null and alternative hypotheses would be 
tested. Explain also why it may be deemed inappropriate to use the 
Wilcoxon rank sum test in this case.
(5)
(ii) Carry out the Wilcoxon signed-rank test at the 5% level of significance, 
and report your conclusions.
(6)