---
title: "Probability Distributions with R"
subtitle: "Introduction to R"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


```{r echo=FALSE}
# - HC Module 3 2011
# - RSS 2011
# - EXAMINATIONS OF THE HONG KONG STATISTICAL SOCIETY
# - HIGHER CERTIFICATE IN STATISTICS, 2011
# - MODULE 3 : Basic statistical methods
# - Question 1
```

1. Ten plants of a certain type, grown under standard conditions and treated with a new 
brand of fertiliser, attain the following heights (in cm).
25 28 24 23 27 30 24 21 28 30
The sample mean is 26.0 and the sample variance is 9.33. From extensive previous 
experiments, it is known that plants of the same type, grown under similar conditions 
but treated with a standard fertiliser, attain a mean height of 25.0 cm.
(i) Write down a statistical model for the distribution of plant heights in the 
population from which the sample is assumed to be drawn, including the 
definition of any unknown parameters.
(3)
(ii) Obtain a 99% confidence interval for the population mean height of plants 
treated with the new fertiliser.
(4)
(iii) Carefully specifying your hypotheses, conduct a test at the 10% significance 
level to decide whether there is evidence to suggest that the new fertiliser 
produces plants of greater mean height than the standard fertiliser. State your 
conclusions.
(10)
(iv) Provide a 95% confidence interval for the population variance of the heights of 
plants treated with the new brand of fertiliser.
(3)
3
