---
title: "Probability Distributions with R"
subtitle: "Introduction to R"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


```{r echo=FALSE}
# - HC Module 3 2012
# - RSS 2012
# - EXAMINATIONS OF THE HONG KONG STATISTICAL SOCIETY
# - HIGHER CERTIFICATE IN STATISTICS, 2012
# - MODULE 3 : Basic statistical methods
# - Question 1
```

1. The following are the scores on an intelligence test (the IQ) of a sample of 10 students 
at a large school.
$$\{107 99 101 93 99 103 134 132 103 109\}$$

Three confidence intervals for the mean IQ for the population of all the students at the 
school have been calculated using a computer package. The output is shown below.
Variable N Mean StDev SE Mean 95% CI
IQ 10 108.000 13.904 4.397 (98.053, 117.947)
Variable N Mean StDev SE Mean ? CI
IQ 10 108.000 13.904 4.397 (99.940, 116.060)
Variable N Mean StDev SE Mean ? CI
IQ 10 108.000 13.904 4.397 (93.711, 122.289)

---
#### Exercises 
(i) The first interval is a 95% confidence interval. One of the second and third is a 
90% confidence interval and the other is a 99% confidence interval. Which is 
which, and how can you tell?

(ii) What assumptions have been made in computing the intervals? What 
distribution has been used in calculating the intervals, and why?
(5)
(iii) Explain the meaning of the above 95% confidence interval as if to a nonstatistician.
(3)
(iv) Explain how the standard deviation is computed and how the standard error of 
the mean is computed.
(3)
(v) Let be the population IQ standard deviation. Construct a 95% confidence 
interval for .

---