---
title: "Statistical Analysis with R"
subtitle: "Introduction to R"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

```{r echo=FALSE}
# - HC Module 3 2011
# - RSS 2011
# - EXAMINATIONS OF THE HONG KONG STATISTICAL SOCIETY
# - HIGHER CERTIFICATE IN STATISTICS, 2011
# - MODULE 3 : Basic statistical methods
# - Question 4
```


4. It is commonly assumed in biblical scholarship that the authors of the gospels of 
Matthew and Luke independently of each other made use of the earlier gospel of 
Mark, and it has been suggested that, if this was so, then Matthew tended to follow the 
text of Mark more closely than did Luke. 

To test whether this suggestion is supported 
by statistical evidence, a random sample of 15 sections of text were taken from Mark 
and compared with each of the parallel sections in Matthew and Luke. A numerical 
measure of similarity, which takes values between 0 and 1, was used to construct the 
two variables, "MtSim", which measures the similarity between Matthew and Mark, 
and "LkSim", which measures the similarity between Luke and Mark. The values of 
these two variables are given below for each of the selected sections of text, larger 
values indicating greater similarity.
MtSim LkSim
0.106 0.199
0.499 0.604
0.465 0.382
0.299 0.301
0.472 0.100
0.642 0.529
0.590 0.386
0.233 0.322
0.308 0.094
0.319 0.128
0.402 0.459
0.672 0.252
0.772 0.483
0.408 0.314
0.337 0.293

---
(i) Explain why it may be deemed appropriate to use the Wilcoxon signed-rank 
test for these data and state what null and alternative hypotheses are being 
tested.

(ii) Carry out the Wilcoxon signed-rank test at the 5% significance level and state 
your conclusions.

(iii) As an alternative to the signed-rank test, carry out the sign test at the 5% 
significance level and state your conclusions.

(iv) Comment briefly on why the conclusions differ when the sign test is used 
instead of the signed-rank test.

---