\documentclass[a4paper,12pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{enumerate}
\usepackage{framed}
\usepackage{epsfig}
\usepackage{subfigure}
\usepackage{fancyhdr}

\setcounter{MaxMatrixCols}{10}
%TCIDATA{OutputFilter=LATEX.DLL}
%TCIDATA{Version=5.00.0.2570}
%TCIDATA{<META NAME="SaveForMode"CONTENT="1">}
%TCIDATA{LastRevised=Wednesday, February 23, 201113:24:34}
%TCIDATA{<META NAME="GraphicsSave" CONTENT="32">}
%TCIDATA{Language=American English}

\pagestyle{fancy}
\setmarginsrb{20mm}{0mm}{20mm}{25mm}{12mm}{11mm}{0mm}{11mm}
\lhead{MA4505} \rhead{Kevin O'Brien} \chead{Homework 7  } %\input{tcilatex}

\begin{document}
%================================================================%
\subsection*{Homework 7 - Experimental Design}

	\begin{framed}
		\begin{itemize}
			\item Deadline : Midnight Friday Week 11
			\item Submit by Email
			\item Subject Line : "\textbf{MA405 Homework 7 Submission}" \\ \textit{(Dont not write anything else)}
		\end{itemize}
	\end{framed}
%================================================================%
\subsubsection*{Task1 One Way ANOVA (4 Marks) }
\begin{itemize}
	\item You should each have a data set in the "\texttt{GreenhouseYields}" Folder.
	\item Is there a significant difference in yield between the treatment groups?
	\item Include Post-Hoc Tests in your discussion (just pick one and use that - i.e. Tukey's HSD). 
	\item State which groups are similar. Report and comment on the the Confidence Intervals.
	\begin{framed}
	\begin{verbatim}
	Analyse > Compare Means > One Way ANOVA
	\end{verbatim}
	\end{framed}
\end{itemize}
\bigskip
%=======================================================================%
\subsubsection*{Task 2 Two Way ANOVA (3 Marks) }
\begin{enumerate}[(i)]
	\item You should each have a data set in the "\texttt{FuelEfficiency}" Folder.
	\item The Fixed Factors are: \texttt{CarType} and \texttt{FuelType}.
%	\item Your Data Set should be similar to the data sets presented in Unit 8 (page 180).
\item Your output should comprise two tables: \textbf{\textit{Between-Subjects Factors}} and \textbf{\textit{Tests of Between-Subjects Effects}}.
\end{enumerate}
	\begin{framed}
		\begin{verbatim}
		Analyse > General Linear Model > Univariate
		\end{verbatim}
	\end{framed}
%=======================================================================%
\newpage
\subsubsection*{Task 3 Two Way ANOVA with Replicates (3 Marks)}
\begin{enumerate}[(i)]
\item You should each have a data set in the "\texttt{Butter}" Folder.
	\item The Fixed Factors are: \texttt{Processes} and \texttt{Factories}.
\item Is there any significant difference in ``fat content" between the Processes
\item Is there any significant difference in ``fat content" between the Factories
\item Include Post-Hoc Tests in your discussion. State which groups are similar.
	\item State which groups are similar. Report and comment on the the Confidence Intervals
\end{enumerate}


	\begin{framed}
		\begin{verbatim}
		Analyse > General Linear Model > Univariate
		\end{verbatim}
	\end{framed}
\end{document}

