\documentclass[a4paper,12pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{enumerate}
\usepackage{framed}
\usepackage{epsfig}
\usepackage{subfigure}
\usepackage{fancyhdr}

\setcounter{MaxMatrixCols}{10}
%TCIDATA{OutputFilter=LATEX.DLL}
%TCIDATA{Version=5.00.0.2570}
%TCIDATA{<META NAME="SaveForMode"CONTENT="1">}
%TCIDATA{LastRevised=Wednesday, February 23, 201113:24:34}
%TCIDATA{<META NAME="GraphicsSave" CONTENT="32">}
%TCIDATA{Language=American English}

\pagestyle{fancy}
\setmarginsrb{20mm}{0mm}{20mm}{25mm}{12mm}{11mm}{0mm}{11mm}
\lhead{MA4505} \rhead{Kevin O'Brien} \chead{SPSS Homework 8}  %\input{tcilatex}

\begin{document}
%================================================================%
\subsection*{SPSS Homework 8 - Correlation, Linear Models, Chi Square}
	\begin{framed}
		\begin{itemize}
			\item Deadline : Midnight Friday Week 12
			\item Submit by Email
			\item Subject Line : "\textbf{MA405 Homework 8 Submission}" \\ \textit{(Dont not write anything else)}
		\end{itemize}
	\end{framed}

\subsubsection*{SPSS Homework 8 - Task 1 Correlations (1/2 Mark) }
\begin{itemize}
	\item You should have a data set in the Zip file called \texttt{CorrCoef} Folder.
	
\item You should have 10 variables: X1 to X10
\item For the sake of simplicity, you may assume that the data sets are normally distributed.
\item Give an example of a pair of variables that have
	
	\begin{enumerate}
	\item very strong positive linear relationship 
	\item relatively strong positive linear relationship
	\item weak positive linear relationship
	\item no linear relationship
	\item weak negative linear relationship
	\item relatively strong negative linear relationship
	\item very strong negative linear relationship 
\end{enumerate}
\end{itemize}
\subsubsection*{SPSS Homework 8 - Task 2 Simple Linear Regression (1 Mark)}
\begin{itemize}
	\item You should have a data set in the  \texttt{Mileage} folder.
	\item Mileage is the dependent variable, Weight is the dependent variable.
	\item Fit a linear model and report the regression equation.
	\item Report the model summary and the regression anova table.
\end{itemize}

%-------------------------------------------------------------------------------%

\subsubsection*{SPSS Homework 8 - Task 3 Chi Square (1 Mark) }
 You should have a data set in the \texttt{MarketResearchUpdated} folder.

\begin{enumerate}[(i)]
	\item Cross Tabulation of Sex and AgeGroup. This is the only cross-tabulation that you are to provide.
	\item Chi Square Test Statistic output for Sex and Agegroup
	\item Chi Square Test Statistic output for Sex and FaveBrand
	\item Chi Square Test Statistic output for Agegroup and FaveBrand
	\item Briefly Interpret each of the Chi Square Tests.
\end{enumerate}

%\subsubsection*{SPSS Homework 8- Task 5 non Parametric Procedures (2 Marks) }
%\begin{itemize}
%	\item You should have a data set in the Zip file called \texttt{NonParametric} Folder.
%\end{itemize}
\end{document}