\documentclass[a4paper,12pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{enumerate}
\usepackage{framed}
\usepackage{epsfig}
\usepackage{subfigure}
\usepackage{fancyhdr}

\setcounter{MaxMatrixCols}{10}
%TCIDATA{OutputFilter=LATEX.DLL}
%TCIDATA{Version=5.00.0.2570}
%TCIDATA{<META NAME="SaveForMode"CONTENT="1">}
%TCIDATA{LastRevised=Wednesday, February 23, 201113:24:34}
%TCIDATA{<META NAME="GraphicsSave" CONTENT="32">}
%TCIDATA{Language=American English}

\pagestyle{fancy}
\setmarginsrb{20mm}{0mm}{20mm}{25mm}{12mm}{11mm}{0mm}{11mm}
\lhead{MA4505} \rhead{Kevin O'Brien} \chead{Homework 6  } %\input{tcilatex}

\begin{document}
%================================================================%
\subsection*{Homework 6 - Hypothesis Testing for Two Samples}

	\begin{framed}
		\begin{itemize}
			\item Deadline : Midnight Friday Week 10
			\item Submit by Email
			\item Subject Line : "\textbf{MA405 Homework 6 Submission}" \\ \textit{(Dont not write anything else)}
		\end{itemize}
	\end{framed}
\subsubsection*{Task 1 Testing Normality (3 Marks) }
\begin{itemize}
	\item You should each have a data set in the "\texttt{CheckingNormality}" Zip File.
	
\item You should have 5 numeric variables in your file: \texttt{X1},\texttt{X2},\texttt{X3}, \texttt{X4} and \texttt{X5}.
	
\item What I want you to do is to test each of these variables for Normality.
\item Use Formal hypothesis Tests, Graphical Procedures
\item Also make a determination based on summary statistics ( kurtosis, Skewness )

%\item The two procedures I want you to use are contained in the "basic statistics" Sub Menu of \textit{\textbf{Statistics}}
%\begin{enumerate}
%\item Graphical Summary Reports
%\item The Normality Test (with Normal Probability Plot)
%\end{enumerate}
%
%\item Try out some of the various Tests: Anderson Darling, Ryan Joiner Test etc. Use only one test per each variable.

\item Write a short report outlining your conclusions : 5 Pages or so including the images from SPSS. 
\item You dont have to add too much discussion to this one, just a written statement of your conlcusion for each variable, with some justification.
\end{itemize}
%================================================================%
\bigskip

\subsubsection*{Task 2 Two Sample Testing Procedures (4 Marks)}
\begin{enumerate}[(i)]
	\item You should find your data sets in ``\texttt{Factory}" Zip file
	\item Is there a significant difference in means between factories A and B?\\
	(Hint: Use \texttt{Analyse| Compare Means}, and use FactoryCode to specify the groups: here A is Factory Code 1, B is Factory Code 2)
	\item Is there a significant difference in variances between the two factories?
	\item  \textbf{IGNORE} Is the rate of defective components the same for Factory A as for Factory B? 
	\item Use graphics, (i.e boxplots and histograms), to compare all four factories.
	\\ (hint: Use \texttt{Analyse | Explore}, and use \textbf{\textit{factories}} as a factor )

\end{enumerate}
\bigskip
%================================================================%
\subsubsection*{Task 3 Paired T-test (3 Marks) }
\begin{itemize}
	\item You should each have a data set in the "\texttt{BeforeAfter}" Zip File.
	\item Is there a significant difference between Before and After?

\end{itemize}
\bigskip


\end{document}

