---
title: "Statistics with R"
subtitle: "Modelling Data with R - Linear Regression"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)

#Higher Certificate, Paper II, 2006.  Question 3

```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


A sample of 10 sea bass was caught by a fisheries scientist who then measured their length x (in millimetres) and their weight y (in grams).  

The data are given in the table below. 




|  | 1 | 2 | 3 | 4 | 5 | 6 | 7 | 8 | 9 | 10 |  
|:---:|:---:|:---:|:---:|:---:|:---:|:---:|:---:|:---:|:---:|:---:|
|Length (x) |387|366|329|293|273|268|294|198 |185|169 |
|Weight (y)|720|680|480|330|270|220|380|108 |89| 68 |




```{r}
Length<- c(387 , 366, 329, 293, 273, 268, 294, 198 ,185 , 169 )
Weight<- c(720, 680, 480, 330, 270, 220, 380, 108  , 89 ,  68 )

myDF <- data.frame(Length, Weight)

```

---

#### Exercises

1. Plot the weights of the 10 sea bass (on the y or vertical axis) against the corresponding lengths (on the x or horizontal axis). Does it appear appropriate to fit a straight line to these data? 
 
2. Calculate the least-squares estimates of the parameters $\beta_0$ and $\beta_1$ of the regression line $y  = \beta_0 + \beta_1 x$. 

3. Comment on the appropriateness of the regression line estimated in part 1 as a model for the relationship between the weights and lengths of sea bass. 


---


```{r}
lm.model <- lm(Weight~Length,data=myDF)
```


```{r}
coef(lm.model)

```
---
 
<pre><code>
library(ggplot2)

p <- ggplot(data = myDF, aes(x=Length,y=Weight)) 
p <- p + geom_point( col="blue", size=2.4 )
p <- p + stat_smooth(method="lm", col="red")
p <- p + theme_bw()
p
</code></pre>

---
 
```{r echo=FALSE,warning=FALSE}
library(ggplot2)

p <- ggplot(data = myDF, aes(x=Length,y=Weight)) 
p <- p + geom_point( col="blue", size=2.4 )
p <- p + stat_smooth(method="lm", col="red")
p <- p + theme_bw()
p
```


---

#### Conclusion

* The graph suggests that a linear fit will be reasonable, at least as a first approximation. 

* There may be curvature in the relation of weight and length. 

* The first three points are not fitted at all well. 
 
---
 
<pre><code>
library(ggplot2)

p <- ggplot(data = myDF, aes(x=Length,y=Weight)) 
p <- p + geom_point( col="blue", size=2.4 )
p <- p + stat_smooth(method="lm", col="red")

## loess fit
p <- p + stat_smooth(col="blue",se=FALSE)

p <- p + theme_bw()
p
</code></pre>

 
---
 
```{r echo=FALSE,warning=FALSE}
library(ggplot2)

p <- ggplot(data = myDF, aes(x=Length,y=Weight)) 
p <- p + geom_point( col="blue", size=2.4 )
p <- p + stat_smooth(method="lm", col="red")
p <- p + stat_smooth(col="blue",se=FALSE)
p <- p + theme_bw()
p
``` 

---

### Sums of Squares Identities



$$\begin{aligned}
S_{xx} &= { \displaystyle \sum (x_i^2) - \frac{ \left(\sum(x_i) \right)^2 }{n} } \\
&= 812,594 - \frac{(2,761)^2}{10} \\
&= 49,729.6
\\
\end{aligned}$$



```{r}
X <- Length
sum( (X - mean(X) )^2)
```

```{r}
Y <- Weight
sum((Y - mean(Y) )^2)
```

---

*  The slope estimate $\hat{\beta}_1$ is
$$\hat{\beta}_1 = \frac{ S_{xy} }{S_{xx}}$$

$$\begin{aligned} S_{xy} &= \sum (x_i y_i) - \frac{ \sum(x_i) \sum(y_i)}{n} \\
&= 1,075,861 - \frac{2,762\times 3,345}{10} \\
&= 151,972.0\\
\end{aligned}$$

```{r}
# Number of Observations
n <- nrow(myDF)

sum(X*Y) - ( sum(X)*sum(Y))/n
```

---

```{r}
summary(myDF)
```

```{r}
Xbar <- mean(X)
Xbar
```

```{r}
Ybar <- mean(Y)
Ybar
```

---

   
        
*  The slope estimate $\hat{\beta}_1$ is therefore
$$ \hat{\beta}_1= \frac{151,972.0}{49,729.6} = 3.056$$


*  The intercept estimate $\hat{\beta}_0$ 
$$\hat{\beta}_0 = \bar{y} - \hat{\beta}_1 \bar{x}  = 334.5 - (3.056\times 276.2) = -509.56$$

---

```{r}
lm.model <- lm(Weight~Length,data=myDF)
summary(lm.model)

```

---

```{r}
coef(lm.model)

```





---

### Exercise 3

Calculate and interpret the coefficient of determination. 
 
The coefficient of determination is given by 
$$R^2 = \frac{S_{xy}^2}{S_{xx}\times S_{yy}}.$$ 


$$  S_{yy} = 1610009 - \frac{23345^2}{10} =  491,106.5 $$
 
$$  R^2 = \frac{151972.0}{49729.6 \times 491106.5} =  0.9457  $$
 
Thus 94.6% of the total variation in the weights of the sea bass is explained by a linear relationship with their lengths. 


---

#### Correlation Coefficient

```{r}

cor(Length,Weight)


cor(Length,Weight)^2
```

#### R-Square (unadjusted)

```{r}
library(modelr)

rsquare(lm.model,data = myDF)

```

---

### Exercise 3

*  Note that the intercept is $–509.56$, whereas it looks as if it should be much nearer 0 if these are to be fitted.  
*  But the remaining points are fitted reasonably well. 
*  It would however be sensible to examine also a quadratic relationship. 
 
---

---

---
 