---
title: "Probability Distributions with R"
subtitle: "Introduction to R"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```
It is estimated by a particular bank that 25% of credit card customers pay only the minimum amount due on their monthly credit card bill and do not pay the total amount due. 500 credit card customers are randomly selected.


#### Exercises

1. What is the probability that 120 or more of the selected customers pay only the minimum amount due?
2. What is the probability that no more than 90 of the selected customers pay only the minimum amount due?
3. What is the probability that more than 100 but less than 160 of the selected customers pay only the minimum amount due?

---

Binomial Probability Distribution

* $X$ is the number of credit card customers that pay only the minimum amount from the selected sample of 500 customers.
* $n$ is the number of independent trials. Here $n=500$.
* $p$ is the probability of a success. Here a success is the event that the credit card customers pay only the minimum amount. $p=0.25$


$$X \sim \mbox{Bin}(n=500,p=0.25)$$

---

### Exercise 1

What is the probability that 120 or more of the selected customers pay only the minimum amount due?

$$P(X \geq 120) = 1- P(X\leq 119)$$
```{r}

1 - pbinom(119,size=500,prob=0.25)

```


```{r}

pbinom(119,size=500,prob=0.25,lower=FALSE)

```

---

### Exercise 2

What is the probability that no more than 90 of the selected customers pay only the minimum amount due?

$$P(X\leq 90)$$
```{r}

pbinom(90,size=500,prob=0.25)

```

```{r}

# Argument Matching
pbinom(90,500,0.25)

```

---


### Exercise 3

What is the probability that more than 100 but less than 160 of the selected customers pay only the minimum amount due?

Remark - sample points between 101 and 159

```{r}

pbinom(159,size=500,prob=0.25)- pbinom(100,size=500,prob=0.25)

```

---

#### Median of $X$

```{r}

qbinom(0.5,size=500,prob=0.25)

```
  
####  First Quartile
  
```{r}

qbinom(0.25,size=500,prob=0.25)

```
  
####  Third Quartile
  
```{r}

qbinom(0.75,size=500,prob=0.25)

```  


---

---