---
title: "Probability Distributions with R"
subtitle: "Introduction to R"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```
---

Telephone calls arrive at a switchboard at the rate of 40 per hour.  
Assume that the telecentre operators take 3 minutes to deal with a customer query.  Calculate the probability of :
### Exercises 
(a)  2 or more calls arriving in any 3 minute period.
(b)  No phone calls arriving in a 3 minute period,
(c)  Exactly one phone call arriving in any 3 minute period,
(d) What is the average and standard deviation of the number of phone calls arriving in a 3 minute period

---

### Exercise 1

```{r}

dpois(0:4,lambda=2)

```
```{r}
ppois(1,lambda=2,lower.tail = FALSE)

```
```{r}
1- ppois(1,lambda=2)

```

---

### Exercise 2



```{r}

dpois(0:4,lambda=2)

```

---

### Exercise 3



```{r}

dpois(0:4,lambda=2)

```