---
title: "Probability Distributions with R"
subtitle: "Introduction to R"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


```{r echo=FALSE}
# - HC Module 3 2012
# - RSS 2012
# - EXAMINATIONS OF THE HONG KONG STATISTICAL SOCIETY
# - HIGHER CERTIFICATE IN STATISTICS, 2012
# - MODULE 3 : Basic statistical methods
# - Question 3
```


3. The Managing Director (MD) of a company of auditors has to select a team of five 
clerical officers for each major audit. He suspects that he may be biased towards lefthanded writers and decides to have this suspicion tested statistically. He knows that 
20% of his large team of clerical officers are left-handed. His analysis of his last 40 
team selections shows the following distribution of the numbers of selected lefthanders.
Number of left-handers 0 1 2 3 4 5
Frequency 9 13 9 5 2 2
You are to decide whether or not these figures provide evidence of bias.

---
#### Exercises 

(i) Assuming no bias towards or against left-handers, confirm that the expected 
number of times out of 40 that the MD would choose each of 0, 1, 2, 3, 4 or 5 
left-handers in his team is 13.11, 16.38, 8.19, 2.05, 0.26 and 0.01 respectively.
(6)
(ii) Carry out a 2
goodness-of-fit test at the 5% significance level to test whether 
or not the observed distribution might reasonably occur if the MD were not 
biased. State your hypotheses and your conclusions.
(10)
(iii) Discuss the implications of your test result and the data with respect to the 
MD's suspicions.
(4)
