---
title: "Statistics with R"
subtitle: "Modelling Data with R - Correlation"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)

# RW-020016

## Higher Certificate, Module4 - 2013. Question 3

```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


A manufacturer of luxury cosmetics has recently put a new product on the market. 

This product is initially being offered at a wide range of prices, and the company has 
made a survey of its sales y (in 100s) and prices x (in £) across a random sample of 
stores in which it is sold. 

It wishes to examine whether, on the whole, increased price 
is associated with reduced sales. The results are shown in the following table.

||||||||||||
|:---|:---:|:---:|:---:|:---:|:---:|:---:|:---:|:---:|:---:|:---:|:---:|
|Store  |1  |2  |3  |4  |5  |6  |7  |8  |9  |10|
|Price  x  (£)  |27  |30  |37  |47  |55  |62  |70  |80  |95  |99|
|Sales  y  (100s)  |110  |79  |69  |48  |51  |44  |29  |32  |26  |30|
---

#### Exercise 1

Plot the data on a scatter diagram and comment on the relationship, if any, 
between x and y.


```{r}

Price <- c(27, 30, 37, 47, 55, 62, 70, 80, 95, 99)
Sales <- c(110, 79, 69, 48, 51, 44, 29, 32, 26, 30)

Survey <- data.frame(Price,Sales)
head(Survey)
```
---

<pre><code>
library(ggplot2)

p <- ggplot(data=Survey, aes(x=Price,y=Sales))
p <- p + geom_point(pch=18,cex=4,col="black")
p <- p + theme_bw()
p
</code></pre>
---


```{r echo=FALSE, warning=FALSE}
library(ggplot2)

p <- ggplot(data=Survey, aes(x=Price,y=Sales))
p <- p + geom_point(pch=18,cex=4,col="black")
p <- p + theme_bw()
p
```
---

<pre><code>
p <- ggplot(data=Survey, aes(x=Price,y=Sales))
p <- p + geom_point(pch=18,cex=4,col="black")
p <- p + theme_bw()
p <- p + stat_smooth(col="red", se=FALSE)
p
</code></pre>
---


```{r echo=FALSE, warning=FALSE}
p <- ggplot(data=Survey, aes(x=Price,y=Sales))
p <- p + geom_point(pch=18,cex=4,col="black")
p <- p + theme_bw()
p <- p + stat_smooth(col="red", se=FALSE)
p
```
---

#### Exercise 2

A research assistant suggests calculating the product-moment correlation 
coefficient, r, between sales and prices. 

Carry out this calculation and test atthe 1% significance level the null hypothesis of zero correlation against an 
appropriate one-sided alternative. 

```{r}
cor(Sales,Price)
```

---

#### Pearson Correlation


```{r}
cor.test(Sales,Price,data=Survey,
         conf.level=0.99, 
         alternative="less")
```


---



#### Exercise 3

A statistician in the market research department suggests calculating instead 
Spearman's rank correlation coefficient, $r_s$. 

Calculate $r_s$ for these data, and test at the 1% level the null hypothesis of no association between prices and sales 
against an appropriate two-sided alternative.

---

```{r}
cor.test(Sales,Price,data=Survey,conf.level=0.99, method="spearman")
```

---


#### Exercise 4

Comment on the tests used in exercises 2 and 3, stating with a reason which 
you prefer.

#### Comparison of Methods

* In statistics, Spearman's rank correlation coefficient or Spearman's ${\displaystyle \rho }$, named after Charles Spearman and often denoted by the Greek letter ${\displaystyle \rho }$  (rho) or as ${\displaystyle r_{s}}$, is a nonparametric measure of rank correlation (statistical dependence between the rankings of two variables).

* It assesses how well the relationship between two variables can be described using a **monotonic function**.


* Monotonic variables increase (or decrease) in the same direction, but not always at the same rate.
Linear variables increase (or decrease) in the same direction at the same rate.

* The Spearman correlation between two variables is equal to the Pearson correlation between the rank values of those two variables; while Pearson's correlation assesses linear relationships, Spearman's correlation assesses monotonic relationships (whether linear or not). 

---


* If there are no repeated data values, a perfect Spearman correlation of +1 or −1 occurs when each of the variables is a perfect monotone function of the other.

---


---
