---
title: "Probability Distributions with R"
subtitle: "Introduction to R"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

Assume that the lifetime, measured in hours, of a packaged magnetic disk exposed to corrosive gases has a Weibull distribution with $\alpha = 0.4$ and $\beta = 400$.

#### Exercises 

Calculate the probability that
1. a disk lasts at least 700 hours,
2. a disk fails between 200 hours and 600 hours.
3. the median life time
4. the inter-quartile range of the lift times

---

### Weibull Distribution

* $\alpha$ is the shape parameter
* $\beta$ is the scale parameter


#### Details on R implementation
The Weibull distribution with shape parameter ***a*** and scale parameter ***b*** has density given by

$$f(x) = (a/b) (x/b)^{(a-1)} exp(- (x/b)^a)$$

for x > 0. The cumulative distribution function is $$F(x) = 1 - exp(- (x/b)^a)$$ on x > 0, the mean is $$E(X) = b \Gamma(1 + 1/a),$$ and the $$Var(X) = b^2 * (\Gamma(1 + 2/a) - (\Gamma(1 + 1/a))^2)$$.

---


#### Exercise 1

$P(X \geq 700)$

```{r} 


pweibull(700, shape=0.4, scale = 400,lower.tail=FALSE)

```


```{r} 


1 - pweibull(700, shape=0.4, scale = 400)

```


---


#### Exercise 2


$P(X \leq 600)$


```{r}
## Remark
pweibull(600, shape=0.4, scale = 400)

```

$P(X \leq 200)$

```{r}
## Remark
pweibull(200, shape=0.4, scale = 400)

```



---


#### Exercise 2


$P(200 \leq X \leq 600)$

```{r} 

pweibull(600, shape=0.4, scale = 400) - 
  pweibull(200, shape=0.4, scale = 400)

```


---

#### Exercise 3

#### Calculate the median

```{r} 

qweibull(0.5, shape=0.4, scale = 400,lower.tail=FALSE)

```

---

#### Exercise 4

Calculate the Inter-quartile Range 

```{r} 
Q1 <- qweibull(0.25, shape=0.4, scale = 400)


Q3 <-qweibull(0.75, shape=0.4, scale = 400)

```

$$ IQR = Q3-Q1 $$

```{r}
Q3-Q1
```

---


---
