---
title: "Probability Distributions with R"
subtitle: "Introduction to R"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


```{r echo=FALSE}
# - HC Module 3 2011
# - RSS 2011
# - EXAMINATIONS OF THE HONG KONG STATISTICAL SOCIETY
# - HIGHER CERTIFICATE IN STATISTICS, 2011
# - MODULE 3 : Basic statistical methods
# - Question 3
```

3. The data below give the daily numbers of homicides in London for the 1095 days from 
1 April 2004 to 31 March 2007.
Daily number of homicides Frequency
0 713
1 299
2 66
3 16
4 1
≥ 5 0
(i) Calculate, to 4 decimal places, the mean number of homicides per day.
(2)
(ii) If homicides are assumed to happen as random events, then the daily numbers 
of homicides should follow a Poisson distribution. Under the hypothesis that 
the data are a random sample from a Poisson distribution, calculate the 
expected frequencies that correspond to the observed frequencies in the table 
above.
(6)
(iii) Carry out a formal test at the 5% significance level to investigate whether or 
not homicides occurred as random events.
(8)
(iv) For a random sample of 139 homicides where the method of killing was 
known, 29 of the homicides were by shooting. Find a 95% confidence interval 
for the underlying population proportion of homicides that are by shooting.
