---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


For a large portfolio of insurance policies, the underlying distribution for losses in the current year has a Pareto Type II distribution with shape parameter $\alpha=2.9$ and scale parameter $\theta=12.5$. 

All losses in the next year are expected to increases by 5\%. For the losses in the next year, determine the tail-value-at-risk at the security level 95\%.

---

### Analytical Solution

---

* set the seed to 1234
* These two distributions have equal weights, so lets simulate 30000 (thirty thousand) variates for each.

```{r}
set.seed(1234)

# install.package("agop")
library(agop)

```

---

```{r}
N <- 30000

X <- rpareto2(N,k=2.9,s=12.5)

head(X)

```

---

```{r}


hist(X,breaks=1000)

```





---
%%- https://actuarialmodelingpractice.wordpress.com/2017/10/29/practice-problem-set-4-pareto-distribution/
%%- https://actuarialmodelingpractice.wordpress.com/2017/10/29/pareto-type-i-versus-pareto-type-ii/
---