---
title: "Statistics with R"
subtitle: "Introduction to R for Actuarial Students"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---
	
%%Practice Problem 4I

For a property and casualty insurance company, losses follow a mixture of two Pareto Type II distributions with equal weights, with the first Pareto 
distribution having parameters $\alpha=1$ and $\theta=500$ and the second Pareto distribution having parameters $\alpha=2$ and $\theta=500$. 

Determine the value-at-risk at the security level of 95\%.



---

* set the seed to 1234
* These two distributions have equal weights, so lets simulate 30000 (thirty thousand) variates for each.

```{r}
set.seed(1234)

N1 <- 30000
N2 <- 30000
```

---


agop - Aggregation Operators and Preordered Sets

Tools supporting multi-criteria and group decision making, including variable number of criteria, by means of aggregation operators, spread measures, fuzzy logic connectives, fusion functions, and preordered sets. 

Possible applications include, but are not limited to, quality management, scientometrics, software engineering, etc.

```{r}
library(agop)
```



---
%%- https://actuarialmodelingpractice.wordpress.com/2017/10/29/practice-problem-set-4-pareto-distribution/
%%- https://actuarialmodelingpractice.wordpress.com/2017/10/29/pareto-type-i-versus-pareto-type-ii/
---

%%- https://actuarialmodelingpractice.wordpress.com/2017/10/29/practice-problem-set-4-pareto-distribution/
%%- https://actuarialmodelingpractice.wordpress.com/2017/10/29/pareto-type-i-versus-pareto-type-ii/

For a large portfolio of insurance policies, the underlying distribution for losses in the current year has a Pareto Type II distribution with shape parameter $\alpha=2.9$ and scale parameter $\theta=12.5$. 

All losses in the next year are expected to increases by 5\%. For the losses in the next year, determine the tail-value-at-risk at the security level 95\%.

---

* set the seed to 1234
* These two distributions have equal weights, so lets simulate 30000 (thirty thousand) variates for each.

```{r}
set.seed(1234)
```

---
	
%%Practice Problem 4I
For a property and casualty insurance company, losses follow a mixture of two Pareto Type II distributions with equal weights, with the first Pareto 
distribution having parameters $\alpha=1$ and $\theta=500$ and the second Pareto distribution having parameters $\alpha=2$ and $\theta=500$. 
\medskip
Determine the value-at-risk at the security level of 95\%.
