---
title: "Probability Distributions with R"
subtitle: "Introduction to R"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

Assume that the diameter of a critical component is normally distributed with a Mean of 50mm and a Standard Deviation of 2mm.

NB You must draw a rough sketch of the normal curve and estimate the approximate probability of the following measurements occurring on an individual component.
--- 
 i.Between 50 and 51.2mm
 ii.Less than 48.5 mm
 iii.Between 48.2 and 51.2 mm


Use the normal tables to get the exact probabilities for the above.



\section*{Solution}
\X \sim N(\mu,\sigma^2) \

 \textbf{Compute the Z-Score}
\ z_{o} = \frac{ x_{o} - \mu}{\sigma} \


---

### Exercise 1


```{r}


```

---

### Exercise 1


```{r}


```

---

### Exercise 1


```{r}


```