---
title: "Probability Distributions with R"
subtitle: "Introduction to R"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```


The lifetime $X$ (in hundreds of hours) of a certain type of industrial vacuum tube has a Weibull distribution with parameters $\alpha = 2.5$ and $\beta = 200$. 


Using a seed of 1234, simulate 10,000 randomly generated lifetimes, and undertake the following exercises.

#### Exercises

Estimate the following:

1. $E(X)$ and $Var(X)$
2. $P(X\leq 200)$
3. $P(180\leq X \leq 250)$
4. $P(X\geq 300)$.

---

```{r}

lifetimes <- rweibull(10000, shape = 2.5, scale =200)

head(lifetimes)

```

```{r}
fivenum(lifetimes)
```

```{r}
summary(lifetimes)
```


---

### Data Visualization

```{r}
library(ggplot2)
lifetimes_df <- data.frame(lifetimes)


```


---

```{r}

hist(lifetimes, 
     breaks=100,
     col = c("lightblue","lightpink"))

```



---

<pre><code>
p <- ggplot( data=lifetimes_df , aes(x = lifetimes) )
p <- p + geom_histogram(binwidth=10,
                        add = "mean", 
                        alpha=0.75,
                        fill = "lightblue")
                      
p <- p + theme_bw()
p
</code></pre>

---

```{r echo=FALSE,warning=FALSE}
p <- ggplot( data=lifetimes_df , aes(x = lifetimes) )
p <- p + geom_histogram(binwidth=10,
                        add = "mean", 
                        alpha=0.75,
                        fill = "lightblue")
                      
p <- p + theme_bw()
p
```

---


### Density Plot

<pre><code>
p <- ggplot( data=lifetimes_df , aes(x = lifetimes) )
p <- p + geom_density(fill = "lightblue")
p <- p + theme_bw()
p
</code></pre>

---

```{r echo=FALSE,warning=FALSE}
p <- ggplot( data=lifetimes_df , aes(x = lifetimes) )
p <- p + geom_density(fill = "lightpink")
p <- p + theme_bw()
p
```

---

#### Exercise 1 - Mean and Variance

```{r}
mean(lifetimes)

```



```{r}
var(lifetimes)

```

---

#### Exercise 2 - $P(X\leq 200)$

```{r}
mean(lifetimes<=200)

```

---
#### Exercise 3 - $P(180\leq X \leq 250)$



```{r}
# library(dplyr)

mean(between(lifetimes,180,250))

```

---

#### Exercise 4 - $P(X\geq 300)$.

```{r}
mean(lifetimes>=300)
```

---
