---
title: "Probability Distributions with R"
subtitle: "Introduction to R"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

```{css, echo=FALSE}
pre {
  background: #ADD8E6;
  max-width: 100%;
  overflow-x: scroll;
}
```

A machine is used to package bags of potato chips.  Records of the packaging machine indicate that its fill weights are normally distributed with a mean of 455 grams per bag and a standard deviation of 10 grams.

\begin{enumerate}
What proportion of bags filled by this machine will contain more than 470 grams in the long run?
What proportion of bags filled by this machine will contain less than 445 grams in the long run?
 What proportion of bags filled by this machine will be between 465 grams and 475 grams in the long run?
\end{enumerate}

---

### Exercise 1


```{r}


```

---

### Exercise 2


```{r}


```

---

### Exercise 3


```{r}


```