---
title: "olsrr"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(olsrr)
```

Package ‘olsrr’
November 22, 2018
Type Package
Title Tools for Building OLS Regression Models
Version 0.5.2

#### Description 
 Tools designed to make it easier for users, particularly beginner/intermediate R users
to build ordinary least squares regression models. Includes comprehensive regression output,
heteroskedasticity tests, collinearity diagnostics, residual diagnostics, measures of influence,
model fit assessment and variable selection procedures.
Depends R(>= 3.3)
Imports car, checkmate, cli, clisymbols, crayon, dplyr, ggplot2, gh,
glue, goftest, graphics, gridExtra, magrittr, nortest, purrr,
Rcpp, recipes (>= 0.1.4), rlang, shiny, stringr, tibble, tidyr
Suggests caret, covr, descriptr, grid, haven, jsonlite, knitr,
lubridate, readr, readxl, rmarkdown, scales, shinyBS,
shinythemes, testthat, tools, vdiffr
License MIT + file LICENSE
URL https://olsrr.rsquaredacademy.com/,
https://github.com/rsquaredacademy/olsrr
BugReports https://github.com/rsquaredacademy/olsrr/issues
Encoding UTF-8
LazyData true
VignetteBuilder knitr
RoxygenNote 6.1.1
LinkingTo Rcpp
NeedsCompilation yes
Author Aravind Hebbali [aut, cre]
Maintainer Aravind Hebbali <hebbali.aravind@gmail.com>
  Repository CRAN
Date/Publication 2018-11-22 17:30:03 UTC
1
2 R topics documented:
  R topics documented:
  auto . 3
cement 4
fitness 4
hsb . . 4
olsrr . 5
ols_aic 5
ols_apc . . . 6
ols_coll_diag 7
ols_correlations . . . 9
ols_deps . . . 10
ols_fpe 11
ols_hadi . . . 12
ols_hsp . . . 12
ols_launch_app . . . 13
ols_leverage . 14
ols_mallows_cp . . . 14
ols_msep . . 15
ols_plot_added_variable . . 16
ols_plot_comp_plus_resid . 18
ols_plot_cooksd_bar 19
ols_plot_cooksd_chart . . . 20
ols_plot_dfbetas . . . 21
ols_plot_dffits 22
ols_plot_diagnostics 23
ols_plot_hadi 23
ols_plot_obs_fit . . . 24
ols_plot_reg_line . . 25
ols_plot_resid_box . 25
ols_plot_resid_fit . . 26
ols_plot_resid_fit_spread . . 27
ols_plot_resid_hist . 28
ols_plot_resid_lev . . 29
ols_plot_resid_pot . . 29
ols_plot_resid_qq . . 30
ols_plot_resid_regressor . . 31
ols_plot_resid_stand 31
ols_plot_resid_stud . 32
ols_plot_resid_stud_fit . . . 33
ols_plot_response . . 34
ols_pred_rsq . 35
ols_prep_avplot_data 35
ols_prep_cdplot_data 36
ols_prep_cdplot_outliers . . 36
ols_prep_dfbeta_data 37
ols_prep_dfbeta_outliers . . 37
ols_prep_dsrvf_data . 38
auto 3
ols_prep_outlier_obs 38
ols_prep_regress_x . 39
ols_prep_regress_y . 39
ols_prep_rfsplot_fmdata . . 40
ols_prep_rstudlev_data . . . 40
ols_prep_rvsrplot_data . . . 41
ols_prep_srchart_data 41
ols_prep_srplot_data 42
ols_press . . 42
ols_pure_error_anova 43
ols_regress . 45
ols_sbc . . . 47
ols_sbic . . . 48
ols_step_all_possible 49
ols_step_all_possible_betas . 50
ols_step_backward_aic . . . 51
ols_step_backward_p 53
ols_step_best_subset 54
ols_step_both_aic . . 56
ols_step_both_p . . . 57
ols_step_forward_aic 59
ols_step_forward_p . 60
ols_test_bartlett . . . 62
ols_test_breusch_pagan . . . 63
ols_test_correlation . 65
ols_test_f . . 66
ols_test_normality . . 67
ols_test_score 68
ols_update . . 69
rivers 70
rvsr_plot_shiny . . . 70
stepdata . . . 71
surgical . . . 71
Index 72

ols_prep_rfsplot_fmdata
Residual fit spread plot data

#### Description 
 
Data for generating residual fit spread plot.

####Usage 
<pre><code>

ols_prep_rfsplot_fmdata(model)
ols_prep_rfsplot_rsdata(model)

 </code></pre>
#### Arguments 
 
model An object of class lm.

####Examples 
 ```{r} 

model <- lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_prep_rfsplot_fmdata(model)
ols_prep_rfsplot_rsdata(model)
ols_prep_rstudlev_data
Studentized residual vs leverage plot data

#### Description 
 
Generates data for studentized resiudual vs leverage plot.

####Usage 
<pre><code>

ols_prep_rstudlev_data(model)

 <\code><\pre>
#### Arguments 
 
model An object of class lm.
ols_prep_rvsrplot_data 41

####Examples 
 ```{r} 

model <- lm(read ~ write + math + science, data = hsb)
ols_prep_rstudlev_data(model)
ols_prep_rvsrplot_ Set

#### Description 
 
Test Data Set

####Usage 
<pre><code>

rivers

#### Format 

An object of class data.frame with 20 rows and 6 columns.
rvsr_plot_shiny Residual vs regressors plot for shiny app

#### Description 
 
Graph to determine whether we should add a new predictor to the model already containing other
predictors. The residuals from the model is regressed on the new predictor and if the plot shows
non random pattern, you should consider adding the new predictor to the model.

####Usage 
<pre><code>

rvsr_plot_shiny(model, data, variable)

 <\code><\pre>
#### Arguments 
 
model An object of class lm.
data A dataframe or tibble.
variable Character; new predictor to be added to the model.

####Examples 
 ```{r} 

model <- lm(mpg ~ disp + hp + wt, data = mtcars)
rvsr_plot_shiny(model, mtcars, 'drat')
stepdata 71
stepdata Test Data Set

#### Description 
 
Test Data Set

####Usage 
<pre><code>

stepdata

#### Format 

An object of class data.frame with 20000 rows and 7 columns.
surgical Surgical Unit Data Set

#### Description 
 
A dataset containing data about survival of patients undergoing liver operation.

####Usage 
<pre><code>

surgical

#### Format 
function test score
age age, in years
gender indicator variable for gender (0 = male, 1 = female)
alc_mod indicator variable for history of alcohol use (0 = None, 1 = Moderate)
alc_heavy indicator variable for history of alcohol use (0 = None, 1 = Heavy)
y Survival Time
Source
Kutner, MH, Nachtscheim CJ, Neter J and Li W., 2004, Applied Linear Statistical Models (5th
                                                                                        edition). Chicago, IL., McGraw Hill/Irwin.
Index
Topic datasets
auto, 3
cement, 4
fitness, 4
hsb, 4
rivers, 70
stepdata, 71
surgical, 71
auto, 3
cement, 4
fitness, 4
hsb, 4
ols_aic, 5, 7, 11, 13, 15, 16, 47, 49
ols_all_subset (ols_step_all_possible),
49
ols_all_subset_betas
(ols_step_all_possible_betas),
50
ols_apc, 6, 6, 11, 13, 15, 16, 47, 49
ols_avplots (ols_plot_added_variable),
16
ols_bartlett_test (ols_test_bartlett),
62
ols_best_subset (ols_step_best_subset),
54
ols_bp_test (ols_test_breusch_pagan), 63
ols_coll_diag, 7
ols_cooksd_barplot
(ols_plot_cooksd_bar), 19
ols_cooksd_chart
(ols_plot_cooksd_chart), 20
ols_corr_test (ols_test_correlation), 65
ols_correlations, 9
ols_deps, 10
ols_dfbetas_panel (ols_plot_dfbetas), 21
ols_dffits_plot (ols_plot_dffits), 22
ols_diagnostic_panel
(ols_plot_diagnostics), 23
ols_dsrvsp_plot
(ols_plot_resid_stud_fit), 33
ols_eigen_cindex (ols_coll_diag), 7
ols_f_test (ols_test_f), 66
ols_fm_plot
(ols_plot_resid_fit_spread), 27
ols_fpe, 6, 7, 11, 13, 15, 16, 47, 49
ols_hadi, 12, 14, 35, 43
ols_hadi_plot (ols_plot_hadi), 23
ols_hsp, 6, 7, 11, 12, 15, 16, 47, 49
ols_launch_app, 13
ols_leverage, 12, 14, 35, 43
ols_mallows_cp, 6, 7, 11, 13, 14, 16, 47, 49
ols_msep, 6, 7, 11, 13, 15, 15, 47, 49
ols_norm_test (ols_test_normality), 67
ols_ovsp_plot (ols_plot_obs_fit), 24
ols_plot_added_variable, 16
ols_plot_comp_plus_resid, 18
ols_plot_cooksd_bar, 19
ols_plot_cooksd_chart, 20
ols_plot_dfbetas, 21
ols_plot_dffits, 22
ols_plot_diagnostics, 23
ols_plot_fm
(ols_plot_resid_fit_spread), 27
ols_plot_hadi, 23
ols_plot_obs_fit, 24
ols_plot_reg_line, 25
ols_plot_resid_box, 25, 26, 28, 30, 65, 68
ols_plot_resid_fit, 26, 26, 28, 30, 65, 68
ols_plot_resid_fit_spread, 27
ols_plot_resid_hist, 26, 28, 30, 65, 68
ols_plot_resid_lev, 29
ols_plot_resid_pot, 29
ols_plot_resid_qq, 26, 28, 30, 65, 68
ols_plot_resid_regressor, 31
ols_plot_resid_spread
72
INDEX 73
(ols_plot_resid_fit_spread), 27
ols_plot_resid_stand, 31
ols_plot_resid_stud, 32
ols_plot_resid_stud_fit, 33
ols_plot_response, 34
ols_potrsd_plot (ols_plot_resid_pot), 29
ols_pred_rsq, 12, 14, 35, 43
ols_prep_avplot_data, 35
ols_prep_cdplot_data, 36
ols_prep_cdplot_outliers, 36
ols_prep_dfbeta_data, 37
ols_prep_dfbeta_outliers, 37
ols_prep_dsrvf_data, 38
ols_prep_outlier_obs, 38
ols_prep_regress_x, 39
ols_prep_regress_y, 39
ols_prep_rfsplot_fmdata, 40
ols_prep_rfsplot_rsdata
(ols_prep_rfsplot_fmdata), 40
ols_prep_rstudlev_data, 40
ols_prep_rvsrplot_data, 41
ols_prep_srchart_data, 41
ols_prep_srplot_data, 42
ols_press, 12, 14, 35, 42
ols_pure_error_anova, 43
ols_reg_line (ols_plot_reg_line), 25
ols_regress, 45
ols_resp_viz (ols_plot_response), 34
ols_rfs_plot
(ols_plot_resid_fit_spread), 27
ols_rpc_plot
(ols_plot_comp_plus_resid), 18
ols_rsd_boxplot (ols_plot_resid_box), 25
ols_rsd_hist (ols_plot_resid_hist), 28
ols_rsd_plot
(ols_plot_resid_fit_spread), 27
ols_rsd_qqplot (ols_plot_resid_qq), 30
ols_rsdlev_plot (ols_plot_resid_lev), 29
ols_rvsp_plot (ols_plot_resid_fit), 26
ols_rvsr_plot
(ols_plot_resid_regressor), 31
ols_sbc, 6, 7, 11, 13, 15, 16, 47, 49
ols_sbic, 6, 7, 11, 13, 15, 16, 47, 48
ols_score_test (ols_test_score), 68
ols_srsd_chart (ols_plot_resid_stand),
31
ols_srsd_plot (ols_plot_resid_stud), 32
ols_step_all_possible, 49, 52, 54, 55, 57,
60, 61
ols_step_all_possible_betas, 50
ols_step_backward
(ols_step_backward_p), 53
ols_step_backward_aic, 50, 51, 54, 55, 57,
60, 61
ols_step_backward_p, 50, 52, 53, 55, 57, 60,
61
ols_step_best_subset, 50, 52, 54, 54, 57,
60, 61
ols_step_both_aic, 50, 52, 54, 55, 56, 60, 61
ols_step_both_p, 57
ols_step_forward (ols_step_forward_p),
60
ols_step_forward_aic, 50, 52, 54, 55, 57,
59, 61
ols_step_forward_p, 50, 52, 54, 55, 57, 60,
60
ols_stepaic_backward
(ols_step_backward_aic), 51
ols_stepaic_both (ols_step_both_aic), 56
ols_stepaic_forward
(ols_step_forward_aic), 59
ols_stepwise (ols_step_both_p), 57
ols_test_bartlett, 62, 64, 67, 69
ols_test_breusch_pagan, 62, 63, 67, 69
ols_test_correlation, 26, 28, 30, 65, 68
ols_test_f, 62, 64, 66, 69
ols_test_normality, 26, 28, 30, 65, 67
ols_test_score, 62, 64, 67, 68
ols_update, 69
ols_vif_tol (ols_coll_diag), 7
olsrr, 5
olsrr-package (olsrr), 5
plot.ols_step_all_possible
(ols_step_all_possible), 49
plot.ols_step_backward_aic
(ols_step_backward_aic), 51
plot.ols_step_backward_p
(ols_step_backward_p), 53
plot.ols_step_best_subset
(ols_step_best_subset), 54
plot.ols_step_both_aic
(ols_step_both_aic), 56
plot.ols_step_both_p (ols_step_both_p),
57
plot.ols_step_forward_aic
(ols_step_forward_aic), 59
74 INDEX
plot.ols_step_forward_p
(ols_step_forward_p), 60
rivers, 70
rvsr_plot_shiny, 70
stepdata, 71
surgical, 71
