---
title: "``ols_regress``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(olsrr)
```

#### Description 
 
Ordinary least squares regression.

####Usage 
<pre><code>
ols_regress(object, ...)

## S3 method for class 'lm'
ols_regress(object, ...)

</code></pre>

#### Arguments
* ``object``: An object of class "formula" (or one that can be coerced to that class): a symbolic Description of the model to be fitted or class lm.
* ``...``: Other inputs.

####Examples 
 ```{r} 

ols_regress(mpg ~ disp + hp + wt, data = mtcars)
# if model includes interaction terms set iterm to TRUE
ols_regress(mpg ~ disp * wt, data = mtcars, iterm = TRUE)
```


#### Value
``ols_regress`` returns an object of class "ols_regress". An object of class "ols_regress" is a list containing the following components:
* r square root of rsquare, correlation between observed and predicted values of dependent variable
* rsq coefficient of determination or r-square
* adjr adjusted rsquare
* ``sigma``: root mean squared error
* cv coefficient of variation
* mse mean squared error
* mae mean absolute error
* ``aic``: akaike information criteria
* ``sbc``: bayesian information criteria
* ``sbic``: sawa bayesian information criteria
* prsq predicted rsquare
* ``error_df``: residual degrees of freedom
* model_df regression degrees of freedom
* total_df total degrees of freedom
* ess error sum of squares
* rss regression sum of squares
* ``tss``: total sum of squares
* rms regression mean square
* ``ems``: error mean square
* f f statistis
* p p-value for f
* ``n``: number of predictors including intercept
* betas betas; estimated coefficients
* sbetas standardized betas
std_errors standard errors
tvalues t values
pvalues p-value of tvalues
df degrees of freedom of betas
conf_lm confidence intervals for coefficients
title title for the model
dependent character vector; name of the dependent variable
* predictors character vector; name of the predictor variables
* mvars character vector; name of the predictor variables including intercept
* model input model for ols_regress

#### Interaction Terms
If the model includes interaction terms, the standardized betas are computed after scaling and centering
the predictors.
References
https://www.ssc.wisc.edu/~hemken/Stataworkshops/stdBeta/Getting

