---
title: "``ols_press``: PRESS"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(olsrr)
```



#### Description 
 
PRESS (prediction sum of squares) tells you how well the model will predict new data.

#### Usage 
<pre><code>

ols_press(model)
</code></pre>

#### Arguments
model An object of class lm.



####Examples 
```{r} 

model <- lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_press(model)
```

#### Details
The prediction sum of squares (PRESS) is the sum of squares of the prediction error. Each fitted
to obtain the predicted value for the ith observation. Use PRESS to assess your model’s predictive
ability. Usually, the smaller the PRESS value, the better the model’s predictive ability.

#### Value
Predicted sum of squares of the model.
ols_pure_error_anova 43

#### References
Kutner, MH, Nachtscheim CJ, Neter J and Li W., 2004, Applied Linear Statistical Models (5th
                                                                                        edition). Chicago, IL., McGraw Hill/Irwin.
See Also
Other influence measures: ols_hadi, ols_leverage, ols_pred_rsq