---
title: "``ols_step_both_aic``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(olsrr)
```
 Stepwise AIC regression

#### Description 
 
Build regression model from a set of candidate predictor variables by entering and removing predictors based on akaike information criteria, in a stepwise manner until there is no variable left to enter or remove any more.

####Usage 
<pre><code>

ols_step_both_aic(model, details = FALSE)
## S3 method for class 'ols_step_both_aic'
plot(x, ...)
</code></pre>

#### Arguments
* model An object of class lm.
* details Logical; if TRUE, details of variable selection will be printed on screen.
* x An object of class ols_step_both_aic.
* ... Other arguments.


####Examples 
```{r} 
## Not run:
# stepwise regression
model <- lm(y ~ ., data = stepdata)
ols_step_both_aic(model)
# stepwise regression plot
model <- lm(y ~ ., data = stepdata)
k <- ols_step_both_aic(model)
plot(k)
# final model
k$model
## End(Not run)
```

#### Value
ols_step_both_aic returns an object of class "ols_step_both_aic". An object of class "ols_step_both_aic"
is a list containing the following components:
* model model with the least AIC; an object of class lm
* predictors variables added/removed from the model
* method addition/deletion
* aics akaike information criteria
* ess error sum of squares
* rss regression sum of squares
* rsq rsquare
* arsq adjusted rsquare
* steps total number of steps

#### Deprecated Function
ols_stepaic_both() has been deprecated. Instead use ``ols_step_both_aic()``.

#### References
Venables,W. N. and Ripley, B. D. (2002) Modern Applied Statistics with S. Fourth edition. Springer.
ols_step_both_p 57
See Also
Other variable selection procedures: ols_step_all_possible, ols_step_backward_aic, ols_step_backward_p,
ols_step_best_subset, ols_step_forward_aic, ols_step_forward_p
