---
title: "``ols_step_backward_p``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(olsrr)
```

Stepwise backward regression

#### Description 
 
Build regression model from a set of candidate predictor variables by removing predictors based on p values, in a stepwise manner until there is no variable left to remove any more.

#### Usage 

<pre><code>

ols_step_backward_p(model, ...)
## Default S3 method:
ols_step_backward_p(model, prem = 0.3, details = FALSE, ...)
## S3 method for class 'ols_step_backward_p'
plot(x, model = NA, ...)

</code></pre>

#### Arguments
* ``model``: An object of class lm; the model should include all candidate predictor variables.
* ``...``: Other inputs.
* ``prem``: p value; variables with p more than prem will be removed from the model.
* ``details``: Logical; if TRUE, will print the regression result at each step.
* ``x``: An object of class ``ols_step_backward_p``.

#### Value

ols_step_backward_p returns an object of class "ols_step_backward_p". 

An object of class "ols_step_backward_p" is a list containing the following components:

* ``model``: final model; an object of class lm
* ``steps``: total number of steps
* ``removed``: variables removed from the model
* ``rsquare``: coefficient of determination
* ``aic``: akaike information criteria
* ``sbic``: sawa bayesian information criteria
* ``sbc``: schwarz bayes information criteria
* ``adjr``: adjusted r-square
* ``rmse``: root mean square error
* ``mallows_cp``: mallow's Cp
* ``indvar``: predictors




####Examples 

```{r} 

# stepwise backward regression
model <- lm(y ~ ., data = surgical)
ols_step_backward_p(model)

# stepwise backward regression plot
model <- lm(y ~ ., data = surgical)
k <- ols_step_backward_p(model)
plot(k)

# final model
k$model


```

#### Deprecated Function
ols_step_backward() has been deprecated. Instead use ols_step_backward_p().

#### References
Chatterjee, Samprit and Hadi, Ali. Regression Analysis by Example. 5th ed. N.p.: John Wiley &
  Sons, 2012. Print.
See Also
Other variable selection procedures: ols_step_all_possible, ols_step_backward_aic, ols_step_best_subset,
ols_step_both_aic, ols_step_forward_aic, ols_step_forward_p