---
title: "olsrr"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(olsrr)
```

ols_step_best_subset Best subsets regression

#### Description 
 
Select the subset of predictors that do the best at meeting some well-defined objective criterion,
such as having the largest R2 value or the smallest MSE, Mallow’s Cp or AIC.

####Usage 
<pre><code>

ols_step_best_subset(model, ...)
## S3 method for class 'ols_step_best_subset'
plot(x, model = NA, ...)
</code></pre>

#### Arguments
* model An object of class lm.
* ... Other inputs.
* x An object of class ols_step_best_subset.


####Examples 

```{r} 
model <- lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_step_best_subset(model)
# plot
model <- lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
k <- ols_step_best_subset(model)
plot(k)
```


#### Value
ols_step_best_subset returns an object of class "ols_step_best_subset". An object of class
"ols_step_best_subset" is a data frame containing the following components:
*  n model number
* predictors predictors in the model
* rsquare rsquare of the model
* adjr adjusted rsquare of the model
* predrsq predicted rsquare of the model
* cp mallow’s Cp
* aic akaike information criteria
* sbic sawa bayesian information criteria
* sbc schwarz bayes information criteria
* gmsep estimated MSE of prediction, assuming multivariate normality
* jp final prediction error
* pc amemiya prediction criteria
* sp hocking’s Sp

#### Deprecated Function
ols_best_subset() has been deprecated. Instead use ols_step_best_subset().
References
Kutner, MH, Nachtscheim CJ, Neter J and Li W., 2004, Applied Linear Statistical Models (5th
                                                                                        edition). Chicago, IL., McGraw Hill/Irwin.
See Also
Other variable selection procedures: ols_step_all_possible, ols_step_backward_aic, ols_step_backward_p,
ols_step_both_aic, ols_step_forward_aic, ols_step_forward_p
