---
title: "``ols_step_both_p ``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(olsrr)
```

Stepwise regression

#### Description 
 
Build regression model from a set of candidate predictor variables by entering and removing predictors
based on p values, in a stepwise manner until there is no variable left to enter or remove any
more.

#### Usage 
<pre><code>

ols_step_both_p(model, ...)
## Default S3 method:
ols_step_both_p(model, pent = 0.1, prem = 0.3,
                details = FALSE, ...)
## S3 method for class 'ols_step_both_p'
plot(x, model = NA, ...)
</code></pre>

#### Arguments
* model An object of class lm; the model should include all candidate predictor variables.
* ... Other arguments.
* pent p value; variables with p value less than pent will enter into the model.
* prem p value; variables with p more than prem will be removed from the model.
* details Logical; if TRUE, will print the regression result at each step.
* x An object of class ols_step_both_p.

#### Examples 
```{r} 
# stepwise regression
model <- lm(y ~ ., data = surgical)
ols_step_both_p(model)
# stepwise regression plot
model <- lm(y ~ ., data = surgical)
k <- ols_step_both_p(model)
plot(k)
# final model
k$model
```

#### Value
ols_step_both_p returns an object of class "ols_step_both_p". An object of class "ols_step_both_p"
is a list containing the following components:
* model final model; an object of class lm
* orders candidate predictor variables according to the order by which they were added or removed from the model method addition/deletion
* steps total number of steps
* predictors variables retained in the model (after addition)
* rsquare coefficient of determination
* aic akaike information criteria
* sbc bayesian information criteria
* sbic sawa’s bayesian information criteria
* adjr adjusted r-square
* rmse root mean square error
* mallows_cp mallow’s Cp
* indvar predictors

#### Deprecated Function
ols_stepwise() has been deprecated. Instead use ols_step_both_p().

#### References
Chatterjee, Samprit and Hadi, Ali. Regression Analysis by Example. 5th ed. N.p.: John Wiley & Sons, 2012. Print.
