---
title: "``ols_step_all_possible``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(olsrr)
```
ols_step_all_possible All possible regression

#### Description 
 
Fits all regressions involving one regressor, two regressors, three regressors, and so on. It tests all
possible subsets of the set of potential independent variables.

#### Usage 
<pre><code>

ols_step_all_possible(model, ...)
## S3 method for class 'ols_step_all_possible'
plot(x, model = NA, ...)
</code></pre>


####  Arguments
* ``model``: An object of class lm.
* ``...``: Other arguments.
* ``x``: An object of class ``ols_best_subset``.

#### Examples 
```{r} 
model <- lm(mpg ~ disp + hp, data = mtcars)
k <- ols_step_all_possible(model)
k
# plot
plot(k)
```

#### Value
ols_step_all_possible returns an object of class "ols_step_all_possible". An object of
class "ols_step_all_possible" is a data frame containing the following components:
* n model number
* predictors predictors in the model
* rsquare rsquare of the model
* adjr adjusted rsquare of the model
* predrsq predicted rsquare of the model
* ``cp``: mallow’s Cp
* ``aic``: akaike information criteria
* ``sbic``: sawa bayesian information criteria
* ``sbc``: schwarz bayes information criteria
* ``gmsep``:  estimated MSE of prediction, assuming multivariate normality
* ``jp``: final prediction error
* ``pc``: amemiya prediction criteria
* ``sp``: hocking’s Sp

#### Deprecated Function
* ``ols_all_subset()`` has been deprecated. Instead use ``ols_step_all_possible()``.

#### References
Mendenhall William and Sinsich Terry, 2012, A Second Course in Statistics Regression Analysis
(7th edition). Prentice Hall

#### See Also
Other variable selection procedures: ols_step_backward_aic, ols_step_backward_p, ols_step_best_subset,
ols_step_both_aic, ols_step_forward_aic, ols_step_forward_p
