---
title: "``ols_coll_diag``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(olsrr)
```



### Collinearity diagnostics

#### Description 

Variance inflation factor, tolerance, eigenvalues and condition indices.

####Usage 
<pre><code>
  
ols_coll_diag(model)
ols_vif_tol(model)
ols_eigen_cindex(model)

</code></pre>

#### Arguments 
  
  model An object of class lm.



####Examples 
```{r} 

# model
model <- lm(mpg ~ disp + hp + wt + drat, data = mtcars)
# vif and tolerance
ols_vif_tol(model)
# eigenvalues and condition indices
ols_eigen_cindex(model)
# collinearity diagnostics
ols_coll_diag(model)
```


#### Details 

Collinearity implies two variables are near perfect linear combinations of one another. Multicollinearity
involves more than two variables. In the presence of multicollinearity, regression estimates
are unstable and have high standard errors.
Tolerance
Percent of variance in the predictor that cannot be accounted for by other predictors.
Steps to calculate tolerance:
  • Regress the kth predictor on rest of the predictors in the model.
• Compute R2 - the coefficient of determination from the regression in the above step.
• Tolerance = 1 􀀀 R2
Variance Inflation Factor
Variance inflation factors measure the inflation in the variances of the parameter estimates due to
collinearities that exist among the predictors. It is a measure of how much the variance of the
estimated regression coefficient 
k is inflated by the existence of correlation among the predictor
variables in the model. A VIF of 1 means that there is no correlation among the kth predictor and
the remaining predictor variables, and hence the variance of 
k is not inflated at all. The general
rule of thumb is that VIFs exceeding 4 warrant further investigation, while VIFs exceeding 10 are
signs of serious multicollinearity requiring correction.
Steps to calculate VIF:
  • Regress the kth predictor on rest of the predictors in the model.
• Compute R2 - the coefficient of determination from the regression in the above step.
• Tolerance = 1=1 􀀀 R2 = 1=T olerance

#### Condition Index
Most multivariate statistical approaches involve decomposing a correlation matrix into linear combinations
of variables. The linear combinations are chosen so that the first combination has the
largest possible variance (subject to some restrictions), the second combination has the next largest
variance, subject to being uncorrelated with the first, the third has the largest possible variance,
subject to being uncorrelated with the first and second, and so forth. The variance of each of these
linear combinations is called an eigenvalue. Collinearity is spotted by finding 2 or more variables
that have large proportions of variance (.50 or more) that correspond to large condition indices. A
rule of thumb is to label as large those condition indices in the range of 30 or larger.

#### Value
ols_coll_diag returns an object of class "ols_coll_diag". An object of class "ols_coll_diag"
is a list containing the following components:
  vif_t tolerance and variance inflation factors
  eig_cindex eigen values and condition index

#### References
Belsley, D. A., Kuh, E., and Welsch, R. E. (1980). Regression Diagnostics: Identifying Influential
Data and Sources of Collinearity. New York: John Wiley & Sons.
