---
title: "``ols_apc``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(olsrr)
```

Amemiya's prediction criterion

#### Description 
 
Amemiya's prediction error.

####Usage 

<pre><code>
ols_apc(model)
</code></pre>

#### Arguments 
 
* ``model``: An object of class lm.

####Examples 

```{r} 
model1 <- lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_apc(model1)
```
```{r} 
model2 <- lm(mpg ~ cyl + wt, data = mtcars)
ols_apc(model2)
```

#### Details 
 
Amemiya'’'s Prediction Criterion penalizes R-squared more heavily than does adjusted R-squared for each addition degree of freedom used on the right-hand-side of the equation. The higher the better for this criterion.
\[((n + p)=(n 􀀀 p))(1 􀀀 (R2))\]

where n is the sample size, p is the number of predictors including the intercept and $R^2$ is the coefficient of determination.

#### Value
Amemiya’s prediction error of the model.

#### References
Amemiya, T. (1976). Selection of Regressors. Technical Report 225, Stanford University, Stanford,
CA.
Judge, G. G., Griffiths, W. E., Hill, R. C., and Lee, T.-C. (1980). The Theory and Practice of
Econometrics. New York: John Wiley & Sons.

#### See Also
Other model selection criteria: ols_aic, ols_fpe, ols_hsp, ols_mallows_cp, ols_msep, ols_sbc,
ols_sbic
