---
title: "olsrr"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(olsrr)
```

ols_step_forward_aic Stepwise AIC forward regression

#### Description 
 
Build regression model from a set of candidate predictor variables by entering predictors based on
akaike information criterion, in a stepwise manner until there is no variable left to enter any more.

####Usage 
<pre><code>

ols_step_forward_aic(model, ...)
## Default S3 method:
ols_step_forward_aic(model, details = FALSE, ...)
## S3 method for class 'ols_step_forward_aic'
plot(x, ...)
Arguments
model An object of class lm.
... Other arguments.
details Logical; if TRUE, will print the regression result at each step.
x An object of class ols_step_forward_aic.
Value
ols_step_forward_aic returns an object of class "ols_step_forward_aic". An object of class
"ols_step_forward_aic" is a list containing the following components:
  model model with the least AIC; an object of class lm
steps total number of steps
predictors variables added to the model
aics akaike information criteria
ess error sum of squares
rss regression sum of squares
rsq rsquare
arsq adjusted rsquare
Deprecated Function
ols_stepaic_forward() has been deprecated. Instead use ols_step_forward_aic().
60 ols_step_forward_p
References
Venables,W. N. and Ripley, B. D. (2002) Modern Applied Statistics with S. Fourth edition. Springer.
See Also
Other variable selection procedures: ols_step_all_possible, ols_step_backward_aic, ols_step_backward_p,
ols_step_best_subset, ols_step_both_aic, ols_step_forward_p

####Examples 
 ```{r} 

# stepwise forward regression
model <- lm(y ~ ., data = surgical)
ols_step_forward_aic(model)
# stepwise forward regression plot
model <- lm(y ~ ., data = surgical)
k <- ols_step_forward_aic(model)
plot(k)
# final model
k$model
