---
title: "ols_step_forward_p"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(olsrr)
```

Stepwise forward regression

#### Description 
 
Build regression model from a set of candidate predictor variables by entering predictors based on
p values, in a stepwise manner until there is no variable left to enter any more.

####Usage 
<pre><code>

ols_step_forward_p(model, ...)
## Default S3 method:
ols_step_forward_p(model, penter = 0.3, details = FALSE,
                   ...)
## S3 method for class 'ols_step_forward_p'
plot(x, model = NA, ...)
</code></pre>

#### Arguments
* model An object of class lm; the model should include all candidate predictor variables.
* ... Other arguments.
* penter p value; variables with p value less than penter will enter into the model details Logical; if TRUE, will print the regression result at each step.
* x An object of class ols_step_forward_p.

#### Examples 

```{r} 

# stepwise forward regression
model <- lm(y ~ ., data = surgical)
ols_step_forward_p(model)
# stepwise forward regression plot
model <- lm(y ~ ., data = surgical)
k <- ols_step_forward_p(model)
plot(k)
# final model
k$model
```

#### Value
ols_step_forward_p returns an object of class "ols_step_forward_p". 
An object of class "ols_step_forward_p" is a list containing the following components:
* model final model; an object of class lm
* steps number of steps
* predictors variables added to the model
* rsquare coefficient of determination
* aic akaike information criteria
* sbc bayesian information criteria
* sbic sawa’s bayesian information criteria
* adjr adjusted r-square
* rmse root mean square error
* mallows_cp mallow’s Cp
* indvar predictors

#### Deprecated Function
ols_step_forward() has been deprecated. Instead use ols_step_forward_p().

#### References
Chatterjee, Samprit and Hadi, Ali. Regression Analysis by Example. 5th ed. N.p.: John Wiley &
  Sons, 2012. Print.
Kutner, MH, Nachtscheim CJ, Neter J and Li W., 2004, Applied Linear Statistical Models (5th                                                                                edition). Chicago, IL., McGraw Hill/Irwin.
See Also
Other variable selection procedures: ols_step_all_possible, ols_step_backward_aic, ols_step_backward_p,
ols_step_best_subset, ols_step_both_aic, ols_step_forward_aic
