
---
title: "``ols_plot_resid_stud_fit``"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(olsrr)
```

Deleted studentized residual vs fitted values plot

#### Description 

Plot for detecting violation of assumptions about residuals such as non-linearity, constant variances
and outliers. It can also be used to examine model fit.

####Usage 
<pre><code>
  
  ols_plot_resid_stud_fit(model)

</code></pre>
  #### Arguments 
  
*  ``model``: An object of class lm.

#### Details 

Studentized deleted residuals (or externally studentized residuals) is the deleted residual divided by
its estimated standard deviation. Studentized residuals are going to be more effective for detecting
outlying Y observations than standardized residuals. If an observation has an externally studentized
residual that is larger than 2 (in absolute value) we can call it an outlier.
34 ols_plot_response
Value
ols_plot_resid_stud_fit returns a list containing the following components:
  outliers a tibble with observation number, fitted values and deleted studentized residuals
that exceed the threshold for classifying observations as outliers/influential
observations
threshold threshold for classifying an observation as an outlier/influential observation
#### Deprecated Function
ols_dsrvsp_plot() has been deprecated. Instead use ols_plot_resid_stud_fit().
See Also
[ols_plot_resid_lev()], [ols_plot_resid_stand()], [ols_plot_resid_stud()]

####Examples 
```{r} 

model <- lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_plot_resid_stud_fit(model)
