---
title: "``ols_plot_dfbetas``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(olsrr)
```

### DFBETAs panel

#### Description 
 
Panel of plots to detect influential observations using DFBETAs.

####Usage 
<pre><code>

ols_plot_dfbetas(model)

</code></pre>

#### Arguments 
 
* ``model``: An object of class lm.

#### Details 
 
DFBETA measures the difference in each parameter estimate with and without the influential point.
There is a DFBETA for each data point i.e if there are n observations and k variables, there will be
n - k DFBETAs. In general, large values of DFBETAS indicate observations that are influential in
estimating a given parameter. Belsley, Kuh, and Welsch recommend 2 as a general cutoff value to
indicate influential observations and 2=
  p
(n) as a size-adjusted cutoff.


####Examples 
```{r} 

model <- lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_plot_dfbetas(model)
```

#### Value
list; ols_plot_dfbetas returns a list of tibbles (for intercept and each predictor) with the observation
number and DFBETA of observations that exceed the threshold for classifying an observation #
as an outlier/influential observation.

#### Deprecated Function
ols_dfbetas_panel() has been deprecated. Instead use ols_plot_dfbetas().
References
Belsley, David A.; Kuh, Edwin; Welsh, Roy E. (1980). Regression Diagnostics: Identifying Influential
Data and Sources of Collinearity.
Wiley Series in Probability and Mathematical Statistics. New York: John Wiley & Sons. pp. ISBN
0-471-05856-4.

#### See Also 

[ols_plot_dffits()]
