---
title: "``ols_plot_resid_fit``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(olsrr)
```

 Residual vs fitted plot

#### Description 
 
Scatter plot of residuals on the y axis and fitted values on the x axis to detect non-linearity, unequal
error variances, and outliers.

####Usage 
<pre><code>

ols_plot_resid_fit(model)

</code></pre>
#### Arguments 
 
model An object of class lm.

#### Details 
 
Characteristics of a well behaved residual vs fitted plot:
  • The residuals spread randomly around the 0 line indicating that the relationship is linear.
• The residuals form an approximate horizontal band around the 0 line indicating homogeneity
of error variance.
• No one residual is visibly away from the random pattern of the residuals indicating that there
are no outliers.
Deprecated Function
ols_rvsp_plot() has been deprecated. Instead use ols_plot_resid_fit().

#### See Also 

Other residual diagnostics: ols_plot_resid_box, ols_plot_resid_hist, ols_plot_resid_qq,
ols_test_correlation, ols_test_normality
ols_plot_resid_fit_spread 27

####Examples 
```{r} 

model <- lm(mpg ~ disp + hp + wt, data = mtcars)
ols_plot_resid_fit(model)
```
