---
title: "``auto``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(olsrr)
```

ols_plot_resid_fit_spread
Residual fit spread plot

#### Description 
 
Plot to detect non-linearity, influential observations and outliers.

####Usage 
<pre><code>

ols_plot_resid_fit_spread(model)
ols_plot_fm(model)
ols_plot_resid_spread(model)

</code></pre>
#### Arguments 
 
model An object of class lm.

#### Details 
 
Consists of side-by-side quantile plots of the centered fit and the residuals. It shows how much
variation in the data is explained by the fit and how much remains in the residuals. For inappropriate
models, the spread of the residuals in such a plot is often greater than the spread of the centered fit.
Deprecated Function
ols_rfs_plot(), ols_fm_plot() and ols_rsd_plot() has been deprecated. Instead use ols_plot_resid_fit_spread(),
ols_plot_fm() and ols_plot_resid_spread().
References
Cleveland, W. S. (1993). Visualizing Data. Summit, NJ: Hobart Press.

####Examples 
```{r} 

# model
model <- lm(mpg ~ disp + hp + wt, data = mtcars)
# residual fit spread plot
ols_plot_resid_fit_spread(model)
28 ols_plot_resid_hist
# fit mean plot
ols_plot_fm(model)
# residual spread plot
ols_plot_resid_spread(model)
```
