---
title: "``olsrr::ols_plot_cooksd_bar``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(olsrr)
```



 Cooks’ D bar plot

#### Description 
 
Bar Plot of cook’s distance to detect observations that strongly influence fitted values of the model.

####Usage 
<pre><code>

ols_plot_cooksd_bar(model)

</code></pre>
#### Arguments 
 
model An object of class lm.

#### Details 
 
Cook’s distance was introduced by American statistician R Dennis Cook in 1977. It is used to
identify influential data points. It depends on both the residual and leverage i.e it takes it account
both the x value and y value of the observation.
Steps to compute Cook’s distance:
  • Delete observations one at a time.
• Refit the regression model on remaining n 􀀀 1 observations
• examine how much all of the fitted values change when the ith observation is deleted.
A data point having a large cook’s d indicates that the data point strongly influences the fitted values.
Value
ols_plot_cooksd_bar returns a list containing the following components:
  outliers a tibble with observation number and cooks distance that exceed threshold
threshold threshold for classifying an observation as an outlier
Deprecated Function
ols_cooksd_barplot() has been deprecated. Instead use ols_plot_cooksd_bar().

#### See Also 

[ols_plot_cooksd_chart()]

####Examples 
```{r} 

model <- lm(mpg ~ disp + hp + wt, data = mtcars)
ols_plot_cooksd_bar(model)
```
