---
title: "``ols_plot_resid_stud``"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(olsrr)
```
### Studentized residual plot

#### Description 

Graph for identifying outliers.

####Usage 

<pre><code>
ols_plot_resid_stud(model)
</code></pre>

#### Arguments 
  
*  ``model``: An object of class lm.


####Examples 
```{r} 

model <- lm(mpg ~ disp + hp + wt, data = mtcars)
ols_plot_resid_stud(model)
```

#### Details 

Studentized deleted residuals (or externally studentized residuals) is the deleted residual divided by
its estimated standard deviation. Studentized residuals are going to be more effective for detecting
outlying Y observations than standardized residuals. If an observation has an externally studentized
residual that is larger than 3 (in absolute value) we can call it an outlier.

#### Value
ols_plot_resid_stud returns a list containing the following components:
  outliers a tibble with observation number and studentized residuals that exceed
threshold
for classifying an observation as an outlier
threshold threshold for classifying an observation as an outlier
#### Deprecated Function
ols_srsd_plot() has been deprecated. Instead use ols_plot_resid_stud().
See Also
[ols_plot_resid_stand()]
