---
title: "``rockchalk::addLines``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rockchalk)
```


#### Description

addLines superimposes regression lines on a plotted plane. The examples will demonstrate the intended usage.
<pre><code>


addLines(to = NULL, from = NULL, col, lwd = 2, lty = 1)

</code></pre>
#### Arguments

* to a 3d plot object produced by plotPlane
* from output from a plotSlopes or plotCurves function (class="rockchalk")
* col color of plotted lines (default: "red")
* ``lwd``:  line width of added lines (default: 2)
* lty line type of added lines (default: 1)



#### Examples
```{r}


##library(rockchalk)
set.seed(12345)
dfadd <- genCorrelatedData2(100, means = c(0,0,0,0), sds = 1, rho = 0,
beta = c(0.03, 0.01, 0.1, 0.4, -0.1), stde = 2)
dfadd$xcat1 <- gl(2,50, labels=c("M","F"))
dfadd$xcat2 <- cut(rnorm(100), breaks=c(-Inf, 0, 0.4, 0.9, 1, Inf),
labels=c("R", "M", "D", "P", "G"))
dfadd$y2 <- 0.03 + 0.1*dfadd$x1 + 0.1*dfadd$x2 +
0.25*dfadd$x1*dfadd$x2 + 0.4*dfadd$x3 - 0.1*dfadd$x4 +
0.2 * as.numeric(dfadd$xcat1) +
contrasts(dfadd$xcat2)[as.numeric(dfadd$xcat2), ] %*% c(-0.1, 0.1, 0.2, 0) +
1 * rnorm(100)
summarize(dfadd)
## linear ordinary regression
m1 <- lm(y ~ x1 + x2 + x3 + x4, data = dfadd)
summary(m1)
mcDiagnose(m1)
## These will be parallel lines
plotSlopes(m1, plotx = "x1", modx = "x2", modxVals = "std.dev.",
n = 5, main = "A plotSlopes result with \"std.dev.\" values of modx")
m1ps <- plotSlopes(m1, plotx = "x1", modx = "x2", modxVals = c(-2,0,2))
m1pp <- plotPlane(m1, plotx1 = "x1", plotx2 = "x2",
ticktype = "detailed", npp = 10)
addLines(from = m1ps, to = m1pp, lty = 1, lwd = 2)
m1pp <- plotPlane(m1, plotx1 = "x1", plotx2 = "x2", ticktype = "detailed",
npp = 10)
addLines(from = m1ps, to = m1pp, lty = 2, lwd = 5, col = "green")
## Other approach would wrap same into the linesFrom argument in plotPlane
plotPlane(m1, plotx1 = "x1", plotx2 = "x2", ticktype = "detailed",
npp = 10, linesFrom = m1ps)
## Need to think more on whether dotted lines from ps object should
## be converted to solid lines in plotPlane.
```


#### Details

From an educational stand point, the objective is to assist with the student’s conceptualization of
the two and three dimensional regression relationships.
Value
NULL, nothing, nicht, nada.
Author(s)
Paul E. Johnson <pauljohn@ku.edu>
addLines 5
