---
title: "``rockchalk::combineLevels``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rockchalk)
```
 recode a factor by "combining" levels

#### Description

This makes it easy to put levels together and create a new factor variable. If a factor variable is
currently coded with levels c("Male","Female","Man", "M"), and the user needs to combine the
redundant levels for males, this is the function to use! This is a surprisingly difficult problem in R.

#### Usage
<pre><code>


combineLevels(fac, levs, newLabel = "combinedLevels")

</code></pre>
#### Arguments

* fac An R factor variable, either ordered or not.
* levs The levels to be combined. Users may specify either a numerical vector of level values, such as c(1,2,3), to combine the first three elements of level(fac), or they may specify level names. This can be done as a character vector of
*correctly spelled* factor values, such as c("Yes","Maybe","Always") or it may
be provided as a subset of the output from levels, such as levels(fac)[1:3].
* ``newLabel``: A character string that represents the label of the new level to be created when levs values are combined.



#### Examples
```{r}


x <- c("M","A","B","C","A","B","A","M")
x <- factor(x)
levels(x)
x2a <- combineLevels(x, levs = c("M","A"), newLabel = "M_or_A")
addmargins(table(x2a, x, exclude=NULL))
x2b <- combineLevels(x, c(1,4), "M_or_A")
addmargins(table(x2b, x, exclude=NULL))
x3 <- combineLevels(x, levs = c("M","A","C"), newLabel = "MAC")
addmargins(table(x3, x, exclude=NULL))
## Now an ordinal factor
z <- c("M","A","B","C","A","B","A","M")
z <- ordered(z)
levels(z)
table(z, exclude=NULL)
z2a <- combineLevels(z, levs = c(1,2), "Good")
addmargins(table(z2a, z, exclude = NULL))
z2b <- combineLevels(z, levs = c("A","B"), "AorB")
addmargins(table(z2b, z, exclude = NULL))
```

#### Details

If the factor is an ordinal factor, then levels may be combined only if they are adjacent. A factor
with levels c("Lo","Med","Hi","Extreme") allows us to combine responses "Lo" and "Med", while
it will NOT allow us to combine "Lo" with "Hi".
A non-ordered factor can be reorganized to combine any values, no matter what positions they
occupy in the levels vector.

#### Value
A new factor variable, with unused levels removed.
Author(s)
Paul E. Johnson <pauljohn@ku.edu>
