---
title: "``formatSummarizedNumerics``"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rockchalk)
```

Reformat numeric summarize output as one column per variable, similar to R summary

#### Description

The summarizeNumeric function returns a data frame with the variable names on the rows and
summary statistics (mean, median, std. deviation) in the columns.This transposes and abbreviates
the information to look more like R summary.

#### Usage
<pre><code>
formatSummarizedNumerics(x, ...)
</code></pre>

#### Arguments
* x numeric summaries from summarize function
* ... Other arguments, such as digits



#### Examples

```{r}
set.seed(21234)
X <- matrix(rnorm(10000), ncol = 10, dimnames = list(NULL, paste0("xvar", 1:10)))
Xsum <- summarize(X)
Xsum$numerics
formatSummarizedNumerics(Xsum$numerics)
formatSummarizedNumerics(Xsum$numerics, digits = 5)
Xsum.fmt <- formatSummarizedNumerics(Xsum$numerics)
str(Xsum.fmt)
```
Value
An R table object
Author(s)
Paul Johnson
