---
title: "``genCorrelatedData``"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rockchalk)
```



genCorrelatedData Generates a data frame for regression analysis

#### Description

The output is a data frame (x1, x2, y) with user-specified correlation between x1 and x2. The y
(output) variable is created according to the equation
$$y = beta1 + beta2  x1 + beta3  x2 + beta4  x1  x2 + e:$$
  The arguments determine the scales of the X matrix, the random error, and the slope coefficients.

#### Usage
<pre><code>
genCorrelatedData(N = 100, means = c(50, 50), sds = c(10, 10), rho = 0,
                  stde = 1, beta = c(0, 0.2, 0.2, 0))
</code></pre>
  
#### Arguments
* N Number of cases desired
* means 2-vector of means for x1 and x2
* sds 2-vector of standard deviations for x1 and x2
* rho Correlation coefficient for x1 and x2
* stde standard deviation of the error term in the data generating equation
* beta beta vector of at most 4 coefficients for intercept, slopes, and interaction


#### Examples
```{r}
## 1000 observations of uncorrelated x1 and x2 with no
## interaction between x1 and x2
dat <- genCorrelatedData(N=1000, rho=0, beta=c(1, 1.0, -1.1, 0.0))
mcGraph1(dat$x1, dat$x2, dat$y, theta=20, phi=8,
         ticktype="detailed", nticks=10)
m1 <- lm(y ~ x1 + x2, data = dat)
plotPlane(m1, plotx1 = "x1", plotx2 = "x2")
```

#### Details
The vector (x1,x2) is drawn from a multivariate normal distribution in which the expected value
(argument means). The covariance matrix of X is built from the standard deviations (sds) and the
specified correlation between x1 and x2 (rho). It is also necessary to specify the standard deviation
of the error term (stde) and the coefficients of the regression equation (beta).
