---
title: "``rockchalk::getDeltaRsquare``"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rockchalk)
```
 Calculates the delta R-squares, also known as squared semi-partial correlation coefficients.

#### Description

The change in the R-square when a variable is removed from a regression is called delta R-square.
It is sometimes suggested as a way to determine whether a variable has a substantial effect on an
outcome. This is also known as the squared semi-partial correlation coefficient.

#### Usage
<pre><code>
getDeltaRsquare(model)
</code></pre>


#### Arguments
* ``model``: a fitted regression model

#### Examples

```{r}
dat1 <- genCorrelatedData(N=250, means=c(100,100),
     sds=c(30,20), rho=0.0, stde = 7, beta=c(1.1, 2.4, 4.1, 0))
m1 <- lm(y ~ x1 + x2, data=dat1)

getDeltaRsquare(m1)
```

```{r}
## more problematic in presence of collinearity
dat2 <- genCorrelatedData(N=250, means=c(100,100),
       sds=c(30,20), rho=0.6, stde = 7, beta=c(1.1, 2.4, 4.1, 0))
m2 <- lm(y ~ x1 + x2, data=dat2)
getDeltaRsquare(m2)
```



### Value
a vector of estimates of the delta R-squares

#### Author(s)
Paul E. Johnson <pauljohn@ku.edu>
