---
title: "``lazyCov``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rockchalk)
```

lazyCov Create covariance matrix from correlation and standard deviation information

#### Description

This is a flexible function that allows lazy R programmers to create covariance matrix. The user
may be lazy because the correlation and standard deviation infomation may be supplied in a variety
of formats.

#### Usage
<pre><code>
lazyCov(Rho, Sd, d)
</code></pre>

#### Arguments
* ``Rho``: Required. May be a single value (correlation common among all variables), a vector of the lower triangular values (vech) of a correlation matrix, or a symmetric matrix of correlation coefficients.
* ``Sd``: Required. May be a single value (standard deviation common among all variables) or a vector of standard deviations, one for each variable.
* ``d``: Optional. Number of rows or columns. lazyCov may be able to deduce the required dimension of the final matrix from the input. However, when the user supplies only a single value for both Rho and Sd, d is necessary.

#### Value
covariance matrix.
Author(s)
<pauljohn@ku.edu>

#### Examples
```{r}
##correlation 0.8 for all pairs, standard deviation 1.0 of each
lazyCov(Rho = 0.8, Sd = 1.0, d = 3)
## supply a vech (lower triangular values in a column)
lazyCov(Rho = c(0.1, 0.2, 0.3), Sd = 1.0)
## supply vech with different standard deviations
lazyCov(Rho = c(0.1, 0.2, 0.3), Sd = c(1.0, 2.2, 3.3))
newRho <- lazyCor(c(0.5, 0.6, 0.7, -0.1, 0.1, 0.2))
lazyCov(Rho = newRho, Sd = 1.0)
lazyCov(Rho = newRho, Sd = c(3, 4, 5, 6))
```

lmAuxiliary Estimate leave-one-variable-out regressions

#### Description

This is a convenience for analysis of multicollinearity in regression.

#### Usage
<pre><code>
lmAuxiliary(model)
</code></pre>

####  Arguments

model a fitted regression model

####   Value
a list of fitted regressions, one for each omitted variable.

####  Author(s)
Paul E. Johnson <pauljohn@ku.edu>
