---
title: "Rockchalk"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rockchalk)
```

magRange magRange Magnify the range of a variable.

#### Description

By default, R’s range function returns the minimum and maximum values of a variable. This returns
a magnified range. It is used for some plotting functions in the rockchalk package

#### Usage

magRange(x, mult = 1.25)
Arguments
x an R vector variable
mult a multiplier by which to magnify the range of the variable. A value of 1 leaves
the range unchanged. May be a scalar, in which case both ends of the range
are magnified by the same amount. May also be a two valued vector, such as
c(minMag, maxMag), in which case the magnification applied to the minimum
is minMag and the magnification of the maximum is maxMag. After version
1.5.5, mult may be smaller than 1, thus shrinking the range. Setting mult to
values closer to 0 causes the range to shrink to the center point from both sides.
makeSymmetric 41

#### Examples
```{r}
x1 <- c(0, 0.5, 1.0)
range(x1)
magRange(x1, 1.1)
magRange(x1, c(1.1, 1.4))
magRange(x1, 0.5)
magRange(x1, c(0.1, 0.1))
```
```{r}
x1 <- rnorm(100)
range(x1)
magRange(x1)
magRange(x1, 1.5)
magRange(x1, c(1,1.5))
```