---
title: "``Rockchalk::padW0``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Rockchalk)
```


Pad with 0s.

#### Description

Sometimes we receive this c(1, 22, 131) and we need character variables of the same size, such as c("001", "022", "131"). This happens if a user has mistakenly converted a zip code (US regional identifier) like "00231" to a number. This function converts the number back to a 0 padded string.

#### Usage

<pre><code>

padW0(x)

</code></pre>

#### Arguments

* ``x``:   A numeric variable.

#### Examples

```{r}


x <- c(1 , 11, 22, 121, 14141, 31)
(xpad <- padW0(x))
x <- rpois(7, lambda = 11)
(xpad <- padW0(x))
x <- c("Alabama", "Iowa", "Washington")
```
#### Details

This works differently if the number provided is an integer, or a character string. Integers are leftpadded with the character "0". A character string will be left-padded with blanks.

#### Value
A character string vector padded with 0’s

#### Author(s)
Paul Johnson <pauljohn@ku.edu>

