---
title: "``Rockchalk::pctable``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Rockchalk)
``` 
#### Description
Creates a cross tabulation with counts and percentages



#### Usage
<pre><code>


pctable(rv, ...)
## Default S3 method:
pctable(rv, cv, rvlab = NULL, cvlab = NULL,
colpct = TRUE, rowpct = FALSE, rounded = FALSE, ...)
## S3 method for class 'formula'
pctable(formula, data = NULL, rvlab = NULL,
cvlab = NULL, colpct = TRUE, rowpct = FALSE, rounded = FALSE, ...)
## S3 method for class 'character'
pctable(rv, cv, data = NULL, rvlab = NULL,
cvlab = NULL, colpct = TRUE, rowpct = FALSE, rounded = FALSE, ...)

</code></pre>
#### Arguments

rv A row variable name
... Other arguments. So far, the most likely additional 
</code></pre>
#### Arguments
 are to be passed
along to the table function, such as "exclude", "useNA", or "dnn" (which will
override the rvlab and cvlab arguments provided by some methods). Some
methods will also pass along these Arguments
 to model.frame, "subset" "xlev",
"na.action", "drop.unused.levels".
cv Column variable
* rvlab Optional: row variable label
* cvlab Optional: col variable label
* colpct Default TRUE: are column percentags desired in the presentation of this result?
* rowpct Default FALSE: are row percentages desired in the presentation of this result
* rounded Default FALSE, rounds to 10’s for privacy purposes.
* formula A two sided formula.
* data A data frame.

#### Examples
```{r}


dat <- data.frame(x = gl(4, 25),
y = sample(c("A", "B", "C", "D", "E"), 100, replace= TRUE))
pctable(y ~ x, dat)
pctable(y ~ x, dat, exclude = NULL)
pctable(y ~ x, dat, rvlab = "My Outcome Var", cvlab = "My Columns")
pctable(y ~ x, dat, rowpct = TRUE, colpct = FALSE)
pctable(y ~ x, dat, rowpct = TRUE, colpct = TRUE)
pctable(y ~ x, dat, rowpct = TRUE, colpct = TRUE, exclude = NULL)
tab <- pctable(y ~ x, dat, rvlab = "Outcome", cvlab = "Predictor")
dat <- data.frame(x1 = gl(4, 25, labels = c("Good", "Bad", "Ugly", "Indiff")),
x2 = gl(5, 20, labels = c("Denver", "Cincy", "Baltimore", "NY", "LA")),
y = sample(c("A", "B", "C", "D", "E"), 100, replace= TRUE))
tab <- pctable(y ~ x1, data = dat, rvlab = "my row label",
subset = dat$x1 %in% c("Good", "Bad"),
drop.unused.levels = TRUE)
tab <- pctable(y ~ x1, data = dat, rvlab = "my row label",
subset = dat$x1 %in% c("Good", "Bad"))
pctable("y", "x1", dat)
pctable("y", x1, dat)
tab <- pctable(y ~ x2, data = dat, rvlab = "A Giant Variable")
summary(tab, rowpct = TRUE, colpct = FALSE)
tabsum <- summary(tab)
## if user has tables package, can push out to latex or html
if (require(tables) & require(Hmisc)){
tabsumtab <- tables::as.tabular(tabsum)
Hmisc::html(tabsumtab)
fn <- tempfile(pattern = "file", tmpdir = tempdir(),
fileext = ".html")
Hmisc::html(tabsumtab, file = fn)
print(paste("The file saved was named", fn, "go get it."))
if (interactive()) browseURL(fn)
## go get the fn file if you want to import it in document
## Now LaTeX output
## have to escape the percent signs
tabsumtab <- apply(tabsumtab, 1:2, function(x) {gsub("%", "\\\\%", x) })
fn2 <- tempfile(pattern = "file", tmpdir = tempdir(),
fileext = ".tex")
Hmisc::latex(tabsumtab, file = fn2)
perspEmpty 77
print(paste("The file saved was named", fn2, "go get it."))
}
```

This function is pronounced "presentable"! The original purpose was to create a particular kind of cross tabulation that I ask for in class: counts with column percentages. Requests from users have caused a bit more generality to be built into the function. Now, optionally, it will provide row percents. This is a generic function. Most users will find the formula method most convenient. Use
the colpct and rowpct  Arguments  to indicate if column or row percentages are desired.
I suggest most users will use the formula method for this. Running a command like this will, generally, do the right thing:

<pre><code>
tab <- pctable(y ~ x, data = dat)
</code></pre>

There is also a method that will work with characters representing variable names.
tab <- pctable("y", "x", data = dat)

Running the function should write a table in the output console, but it also creates an object (tab).
That object can be displayed in a number of ways. A summary method is provided, so one could look at different representations of the same table.
summary(tab, rowpct = TRUE, colpct = FALSE)
or
summary(tab, rowpct = TRUE, colpct = TRUE)

Tables that include only row or column percentages will be compatible with the html and latex exporters in the excellent tables package.
The method pctable.default is the calculator, I don’t expect many users will need to call it directly.
The formula method is the recommended method for users. Run pctable(myrow ~ mycol, data = dat).
In an earlier version, I gave different advice, so please adjust your usage.
The character method exists only for variety. It accepts character strings rather than a formula to define the columns that should be plotted. The method used most often for most users should be the formula method.

#### Details

Please bear in mind the following. The output object is a list of tables of partial information, which are then assembled in various ways by the print method (print.pctable). A lovely table will appear on the screen, but the thing itself has more information and a less beautiful structure.

A print method is supplied. For any pctable object, it is possible to run follow-ups like print(tab, rowpct = TRUE, colpct = FALSE)
The method print.pctable(tab) assembles the object into (my opinion of) a presentable form.

The print method has argumnets rowpct and colpct that determine which percentages are included in the presentation.
When using character arguments, the row variable rv rowvar must be a quoted string if the user intends the method pctable.character to be dispatched. The column variable cv may be a string or just a variable name (which this method will coerce to a string).

#### Value
A list with tables (count, column percent, row percent) as well as a copy of the call.

#### Author(s)
Paul Johnson <pauljohn@ku.edu>
See Also
tabular and the CrossTable function in gmodels package.

