---
title: "``Rockchalk::perspEmpty``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Rockchalk)
``` 


 perspEmpty

#### Description

Creates a persp plot without drawing anything in the interior. Does equivalent of plot( type="n")
for persp.

#### Usage
<pre><code>


perspEmpty(x1, x2, y, x1lab = "x1", x2lab = "x2", ylab = "y", x1lim,
x2lim, ...)

</code></pre>
#### Arguments

x1 data for the first horizontal axis, an R vector
x2 data for the second horizontal axis, an R vector
y data for the vertical axis, an R vector
x1lab label for the x1 axis, (the one called "xlab" inside persp)
x2lab label for the x2 axis, (the one called "ylab" inside persp)
* ``ylab``:   label for the y (vertical) axis (the one called "zlab" inside persp)
x1lim Optional: limits for x1 axis (should be a vector with 2 elements)
x2lim Optional: limits for x2 axis (should be a vector with 2 elements)
... furtherarguments
 that are passed to persp. Please note Please remember that
y is the vertical axis, but for persp, that is the one I call x2. Thus dot-dotdot

</code></pre>
#### Arguments
 including xlab, ylab, zlab, xlim, ylim, and zlim are going to be
ignored.

#### Details

Regression demonstrations require a blank slate in which points and planes can be drawn. This
function creates that blank persp canvas for those projects. It is not necessary that x1, x2 and y be
vectors of the same length, since this function’s only purpose is to plot an empty box with ranges
determined by the input variables. persp calls the 3 axes x, y, and z, but here they are called x1, x2,
and y.
Value
The perspective matrix that is returned by persp


#### Examples
```{r}
x1 <- 1:10
x2 <- 41:50
y <- rnorm(10)
perspEmpty(x1, x2, y)
res <- perspEmpty(x1, x2, y, ticktype="detailed", nticks=10)
mypoints1 <- trans3d ( x1, x2, y, pmat = res )
points( mypoints1, pch = 16, col= "blue")
```
