---
title: "``predictOMatic``"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rockchalk)
```

 Create predicted values after choosing values of predictors. Can demonstrate marginal effects of the predictor variables.

#### Description

It creates "newdata" frames which are passed to predict. The key idea is that each predictor has certain focal values on which we want to concentrate. We want a more-or-less easy way to spawn complete newdata objects along with fitted values. The newdata function creates those objects, its documentation might be helpful in understanding some nuances.

#### Usage
<pre><code>
predictOMatic(model = NULL, predVals = "margins", divider = "quantile",n = 5, ...)
</code></pre>

#### Arguments

* ``model``: Required. A fitted regression model. A predict method must exist for that model.
* ``predVals``: Optional. How to choose predictor values? Can be as simple as a keyword "auto" or "margins". May also be very fine-grained detail, including 1) a vector of variable names (for which values will be automatically selected) 2) a named
vector of variable names and divider functions, or 3) a list naming variables and values. See details and examples.

* ``divider``: An algorithm name from c("quantile", "std.dev", "seq", "table") or a user-provided function. This sets the method for selecting values of the predictor.  Documentation for the rockchalk methods can be found in the functions cutByQuantile, cutBySD, plotSeq, and cutByTable,.
* n Default = 5. The number of values for which predictions are sought.
* ``...``: Optional arguments to be passed to the predict function. In particular, the Arguments

se.fit and interval are extracted from ... and used to control the output.




#### Examples
```{r}


library(rockchalk)
## Replicate some R classics. The budworm.lg data from predict.glm
## will work properly after re-formatting the information as a data.frame:
## example from Venables and Ripley (2002, pp. 190-2.)
df <- data.frame(ldose = rep(0:5, 2),
sex = factor(rep(c("M", "F"), c(6, 6))),
SF.numdead = c(1, 4, 9, 13, 18, 20, 0, 2, 6, 10, 12, 16))
df$SF.numalive = 20 - df$SF.numdead
budworm.lg <- glm(cbind(SF.numdead, SF.numalive) ~ sex*ldose,
data = df, family = binomial)
predictOMatic(budworm.lg)
predictOMatic(budworm.lg, n = 7)
predictOMatic(budworm.lg, predVals = c("ldose"), n = 7)
predictOMatic(budworm.lg, predVals = c(ldose = "std.dev.", sex = "table"))
```

```{r}
## Now make up a data frame with several numeric and categorical predictors.
set.seed(12345)
N <- 100
x1 <- rpois(N, l = 6)
x2 <- rnorm(N, m = 50, s = 10)
x3 <- rnorm(N)
predictOMatic 99
xcat1 <- gl(2,50, labels = c("M","F"))
xcat2 <- cut(rnorm(N), breaks = c(-Inf, 0, 0.4, 0.9, 1, Inf),
labels = c("R", "M", "D", "P", "G"))
dat <- data.frame(x1, x2, x3, xcat1, xcat2)
rm(x1, x2, x3, xcat1, xcat2)
dat$xcat1n <- with(dat, contrasts(xcat1)[xcat1, , drop = FALSE])
dat$xcat2n <- with(dat, contrasts(xcat2)[xcat2, ])
STDE <- 15
dat$y <- with(dat,
0.03 + 0.8*x1 + 0.1*x2 + 0.7*x3 + xcat1n %*% c(2) +
xcat2n %*% c(0.1,-2,0.3, 0.1) + STDE*rnorm(N))
```

```{r}
## Impose some random missings
dat$x1[sample(N, 5)] <- NA
dat$x2[sample(N, 5)] <- NA
dat$x3[sample(N, 5)] <- NA
dat$xcat2[sample(N, 5)] <- NA
dat$xcat1[sample(N, 5)] <- NA
dat$y[sample(N, 5)] <- NA
summarize(dat)
m0 <- lm(y ~ x1 + x2 + xcat1, data = dat)
summary(m0)
```

```{r}
## The model.data() function in rockchalk creates as near as possible
## the input data frame.
m0.data <- model.data(m0)
summarize(m0.data)
## no predVals: analyzes each variable separately
(m0.p1 <- predictOMatic(m0))
## requests confidence intervals from the predict function
(m0.p2 <- predictOMatic(m0, interval = "confidence"))
## predVals as vector of variable names: gives "mix and match" predictions
(m0.p3 <- predictOMatic(m0, predVals = c("x1", "x2")))
## predVals as vector of variable names: gives "mix and match" predictions
(m0.p3s <- predictOMatic(m0, predVals = c("x1", "x2"), divider = "std.dev."))
```

```{r}
## "seq" is an evenly spaced sequence across the predictor.
(m0.p3q <- predictOMatic(m0, predVals = c("x1", "x2"), divider = "seq"))
(m0.p3i <- predictOMatic(m0, predVals = c("x1", "x2"),
interval = "confidence", n = 3))
```

```{r}
(m0.p3p <- predictOMatic(m0, predVals = c("x1", "x2"), divider = pretty))
## predVals as vector with named divider algorithms.
(m0.p3 <- predictOMatic(m0, predVals = c(x1 = "seq", x2 = "quantile")))
## predVals as named vector of divider algorithms
## same idea, decided to double-check
100 predictOMatic
(m0.p3 <- predictOMatic(m0, predVals = c(x1 = "quantile", x2 = "std.dev.")))
getFocal(m0.data$x2, xvals = "std.dev.", n = 5)
## Change from quantile to standard deviation divider
(m0.p5 <- predictOMatic(m0, divider = "std.dev.", n = 5))
## Still can specify particular values if desired
(m0.p6 <- predictOMatic(m0, predVals = list("x1" = c(6,7),
"xcat1" = levels(m0.data$xcat1))))
(m0.p7 <- predictOMatic(m0, predVals = c(x1 = "quantile", x2 = "std.dev.")))
getFocal(m0.data$x2, xvals = "std.dev.", n = 5)
(m0.p8 <- predictOMatic(m0, predVals = list( x1 = quantile(m0.data$x1,
na.rm = TRUE, probs = c(0, 0.1, 0.5, 0.8,
1.0)), xcat1 = levels(m0.data$xcat1))))
```

```{r}
(m0.p9 <- predictOMatic(m0, predVals = list(x1 = "seq", "xcat1" =
levels(m0.data$xcat1)), n = 8) )
(m0.p10 <- predictOMatic(m0, predVals = list(x1 = "quantile",
"xcat1" = levels(m0.data$xcat1)), n = 5) )
(m0.p11 <- predictOMatic(m0, predVals = c(x1 = "std.dev."), n = 10))
## Previous same as
(m0.p11 <- predictOMatic(m0, predVals = c(x1 = "default"), divider =
"std.dev.", n = 10))
## Previous also same as
(m0.p11 <- predictOMatic(m0, predVals = c("x1"), divider = "std.dev.", n = 10))
(m0.p11 <- predictOMatic(m0, predVals = list(x1 = c(0, 5, 8), x2 = "default"),
divider = "seq"))
m1 <- lm(y ~ log(10+x1) + sin(x2) + x3, data = dat)
m1.data <- model.data(m1)
summarize(m1.data)
(newdata(m1))
(newdata(m1, predVals = list(x1 = c(6, 8, 10))))
(newdata(m1, predVals = list(x1 = c(6, 8, 10), x3 = c(-1,0,1))))
(newdata(m1, predVals = list(x1 = c(6, 8, 10),
x2 = quantile(m1.data$x2, na.rm = TRUE), x3 = c(-1,0,1))))
```

```{r}
(m1.p1 <- predictOMatic(m1, divider = "std.dev", n = 5))
(m1.p2 <- predictOMatic(m1, divider = "quantile", n = 5))
(m1.p3 <- predictOMatic(m1, predVals = list(x1 = c(6, 8, 10),
x2 = median(m1.data$x2, na.rm = TRUE))))
(m1.p4 <- predictOMatic(m1, predVals = list(x1 = c(6, 8, 10),
x2 = quantile(m1.data$x2, na.rm = TRUE))))
(m1.p5 <- predictOMatic(m1))
(m1.p6 <- predictOMatic(m1, divider = "std.dev."))
(m1.p7 <- predictOMatic(m1, divider = "std.dev.", n = 3))
(m1.p8 <- predictOMatic(m1, divider = "std.dev.", interval = "confidence"))
m2 <- lm(y ~ x1 + x2 + x3 + xcat1 + xcat2, data = dat)
## has only columns and rows used in model fit
m2.data <- model.data(m2)
summarize(m2.data)
## Check all the margins
(predictOMatic(m2, interval = "conf"))
## Lets construct predictions the "old fashioned way" for comparison
m2.new1 <- newdata(m2, predVals = list(xcat1 = levels(m2.data$xcat1),
xcat2 = levels(m2.data$xcat2)), n = 5)
predict(m2, newdata = m2.new1)
(m2.p1 <- predictOMatic(m2,
predVals = list(xcat1 = levels(m2.data$xcat1),
xcat2 = levels(m2.data$xcat2)),
xcat2 = c("M","D")))
```

```{r}
## See? same!
## Pick some particular values for focus
m2.new2 <- newdata(m2, predVals = list(x1 = c(1,2,3), xcat2 = c("M","D")))
## Ask for predictions
predict(m2, newdata = m2.new2)
## Compare: predictOMatic generates a newdata frame and predictions in one step
(m2.p2 <- predictOMatic(m2, predVals = list(x1 = c(1,2,3),
xcat2 = c("M","D"))))
(m2.p3 <- predictOMatic(m2, predVals = list(x2 = c(0.25, 1.0),
xcat2 = c("M","D"))))
(m2.p4 <- predictOMatic(m2, predVals = list(x2 = plotSeq(m2.data$x2, 10),
xcat2 = c("M","D"))))
```

```{r}
(m2.p5 <- predictOMatic(m2, predVals = list(x2 = c(0.25, 1.0),
xcat2 = c("M","D")), interval = "conf"))
(m2.p6 <- predictOMatic(m2, predVals = list(x2 = c(49, 51),
xcat2 = levels(m2.data$xcat2),
x1 = plotSeq(dat$x1))))
```

```{r}
plot(y ~ x1, data = m2.data)
by(m2.p6, list(m2.p6$xcat2), function(x) {
lines(x$x1, x$fit, col = x$xcat2, lty = as.numeric(x$xcat2))
})
m2.newdata <- newdata(m2, predVals = list(x2 = c(48, 50, 52),
xcat2 = c("M","D")))
```

```{r}
predict(m2, newdata = m2.newdata)
(m2.p7 <- predictOMatic(m2, predVals = list(x2 = c(48, 50, 52),
xcat2 = c("M","D"))))
(m2.p8 <- predictOMatic(m2,
predVals = list(x2 = range(m2.data$x2, na.rm = TRUE),
xcat2 = c("M","D"))))
```

```{r}
(m2.p9 <- predictOMatic(m2, predVals = list(x2 = plotSeq(m2.data$x2),
x1 = quantile(m2.data$x1, pr =c(0.33, 0.66), na.rm = TRUE),
xcat2 = c("M","D"))))
plot(y ~ x2 , data = m2.data)
by(m2.p9, list(m2.p9$x1, m2.p9$xcat2), function(x) {lines(x$x2, x$fit)})
(predictOMatic(m2, predVals = list(x2 = c(50, 60), xcat2 = c("M","D")),
interval = "conf"))
```

```{r}
## create a dichotomous dependent variable
y2 <- ifelse(rnorm(N) > 0.3, 1, 0)
dat <- cbind(dat, y2)
m3 <- glm(y2 ~ x1 + x2 + x3 + xcat1, data = dat, family = binomial(logit))
summary(m3)
```

```{r}
m3.data <- model.data(m3)
summarize(m3.data)
(m3.p1 <- predictOMatic(m3, divider = "std.dev."))
(m3.p2 <- predictOMatic(m3, predVals = list(x2 = c(40, 50, 60),
xcat1 = c("M","F")),
divider = "std.dev.", interval = "conf"))
```

```{r}
## Want a full accounting for each value of x2?
(m3.p3 <- predictOMatic(m3,
predVals = list(x2 = unique(m3.data$x2),
xcat1 = c("M","F")), interval = "conf"))
## Would like to write a more beautiful print method
## for output object, but don't want to obscure structure from user.
## for (i in names(m3.p1)){
## dns <- cbind(m3.p1[[i]][i], m3.p1[[i]]$fit)
## colnames(dns) <- c(i, "predicted")
## print(dns)
## }


#### Details

If no predVals argument is supplied (same as predVals = "margins", predictOMatic creates a list
of new data frames, one for each predictor variable. It uses the default divider algorithm (see the
divider argument) and it estimates predicted values for n different values of the predictor. A model
with formula y ~ x1 + x2 + x3 will cause 3 separate output data frames, one for each predictor.
They will be named objects in the list.
The default approach will have marginal tables, while the setting predVals = "auto" will create
a single large newdata frame that holds the Cartesian product of the focal values of each predictor.
predVals may be a vector of variable names, or it may be a list of names and particular values.
Whether a vector or a list is supplied, predVals must name only predictors that are fitted
in the model. predictOMatic will choose the mean or mode for variables that are not explicitly
listed, and selected values of the named variables are "mixed and matched" to make a
data set. There are many formats in which it can be supplied. Suppose a regression formula is
y1 ~ sex + income + health + height. The simplest format for predVals will be a vector
of variable names, leaving the selection of detailed values to the default algorithms. For example,
predVals = c("income","height") will cause sex and health to be set at central values
and income and height will have target values selected according to the divider algorithm (see the
argument divider).
The user can spcecify divider algoriths to choose focal values, predvals = c(income = "quantile", height = "std.dev.").
The dividers provided by the rockchalk package are "quantile", "std.dev.", "seq" and "table". Those are discussed more completely in the help for focalVals. The appropriate algorithms will select focal values of the predictors and they will supply n values for each in a "mix and match" data frame. After rockchalk 1.7.2, the divider argument can also be the name of a function, such as R's
pretty.

Finally, users who want very fine grained control over predictOMatic can supply a named list of predictor
values. For example, predVals = list(height = c(5.5, 6.0, 6.5),income = c(10, 20, 30, 40, 50), sex = levels(One can also use algorithm names, predVals = list(height = c(5.5, 6.0, 6.5), income = "quantile")
and so forth. 
The variables named in the predVals argument should be the names of the variables in the raw data
frame, not the names that R creates when it interprets a formula. We want "x", not the transformation
in the functions (not log(x), or as.factor(x) or as.numeric(x)). 

If a formula has a predictor poly(height, 3), then the predVals argument should refer to height, not poly(height, 3). I’ve
invested quite a bit of effort to make sure this "just works" (many alternative packages that calculate
predicted values do not).

It it important to make sure that diagnostic plots and summaries of predictions are calculated with
the exact same data that was used to fit the model. This is surprisingly difficult because formulas
can include things like log(income + d) and so forth. The function model.data is the magic bullet for that part of the problem.
Here is one example sequence that fits a model, discerns some focal values, and then uses predictOMatic.
<pre><code>
d <- 3 alpha <- 13 m1 <- lm(yout ~ xin + xout + poly(xother,2) + log(xercise + alpha), data = dat)
m1dat <- model.data(m1)
</code></pre>
Now, when you are thinking about which values you might like to specify in predVals, use m1dat
to decide. Try
summarize(m1dat)

Then run something like
predictOMatic( m1, predVals = list(xin = median(m1dat$xin), xout = c(1,2,3), xother = quantile(m1dat$xother))

Get the idea?

#### Value
A data frame or a list of data frames.
Author(s)
Paul E. Johnson <pauljohn@ku.edu>
