---
title: "``Rockchalk::rbindFill``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Rockchalk)
```

rbindFill Stack together data frames

#### Description

In the end of the code for plyr::rbind.fill, the author explains that is uses an experimental function to build the data.frame. I would rather not put any weight on an experimental function, so I sat out to create a new rbindFill. This function uses no experimental functions. It does not rely on any functions from packages that are not in base of R, except, of course, for functions in this package.

#### Usage
<pre><code>


rbindFill(...)

</code></pre>
#### Arguments

... Data frames

#### Details

Along the way, I noticed a feature that seems to be a flaw in both rbind and rbind.fill. In the 
#### Examples
```{r}

,
there is a demonstration of the fact that base R rbind and plyr::rbind.fill both have undesirable
properties when data sets containing factors and ordered variables are involved. This function introduces
a "data consistency check" that prevents corruption of variables when data frames are
combined. This "safe" version will notice differences in classes of variables among data.frames and
stop with an error message to alert the user to the problem.
Value
A stacked data frame
Author(s)
Paul Johnson

#### Examples
```{r}


set.seed(123123)
N <- 10000
dat <- genCorrelatedData2(N, means = c(10, 20, 5, 5, 6, 7, 9), sds = 3,
stde = 3, rho = .2, beta = c(1, 1, -1, 0.5))
dat1 <- dat
dat1$xcat1 <- factor(sample(c("a", "b", "c", "d"), N, replace=TRUE))
dat1$xcat2 <- factor(sample(c("M", "F"), N, replace=TRUE),
levels = c("M", "F"), labels = c("Male", "Female"))
dat1$y <- dat$y +
as.vector(contrasts(dat1$xcat1)[dat1$xcat1, ] %*% c(0.1, 0.2, 0.3))
dat1$xchar1 <- rep(letters[1:26], length.out = N)
dat2 <- dat
106 religioncrime
dat1$x3 <- NULL
dat2$x2 <- NULL
dat2$xcat2 <- factor(sample(c("M", "F"), N, replace=TRUE),
levels = c("M", "F"), labels = c("Male", "Female"))
dat2$xcat3 <- factor(sample(c("K1", "K2", "K3", "K4"), N, replace=TRUE))
dat2$xchar1 <- "1"
dat3 <- dat
dat3$x1 <- NULL
dat3$xcat3 <- factor(sample(c("L1", "L2", "L3", "L4"), N, replace=TRUE))
dat.stack <- rbindFill(dat1, dat2, dat3)
str(dat.stack)
## Possible BUG alert about base::rbind and plyr::rbind.fill
## Demonstrate the problem of a same-named variable that is factor in one and
## an ordered variable in the other
dat5 <- data.frame(ds = "5", x1 = rnorm(N),
xcat1 = gl(20, 5, labels = LETTERS[20:1]))
dat6 <- data.frame(ds = "6", x1 = rnorm(N),
xcat1 = gl(20, 5, labels = LETTERS[1:20], ordered = TRUE))
## rbind reduces xcat1 to factor, whether we bind dat5 or dat6 first.
stack1 <- base::rbind(dat5, dat6)
str(stack1)
## note xcat1 levels are ordered T, S, R, Q
stack2 <- base::rbind(dat6, dat5)
str(stack2)
## xcat1 levels are A, B, C, D
## stack3 <- plyr::rbind.fill(dat5, dat6)
## str(stack3)
## xcat1 is a factor with levels T, S, R, Q ...
## stack4 <- plyr::rbind.fill(dat6, dat5)
## str(stack4)
## oops, xcat1 is ordinal with levels A < B < C < D
## stack5 <- rbindFill(dat5, dat6)
