---
title: "``Rockchalk::vech2mat``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Rockchalk)
```

vech2mat Convert a half-vector (vech) into a matrix.

#### Description

Fills a matrix from a vector that represents the lower triangle. If user does not supply a value for
diag, then the vech will fill in the diagonal as well as the strictly lower triangle. If diag is provided
(either a number or a vector), then vech is for the strictly lower triangular part. The default value
for lowerOnly is FALSE, which means that a symmetric matrix will be created. See 
#### Examples
```{r}

 for
a demonstration of how to fill in the lower triangle and leave the diagonal and the upper triangle
empty.

#### Usage
<pre><code>


vech2mat(vech, diag = NULL, lowerOnly = FALSE)

</code></pre>
#### Arguments

vech A vector
diag Optional. A single value or a vector for the diagonal. A vech is a strictly lower
triangluar vech, it does not include diagonal values. diag can be either a single
value (to replace all elements along the diagonal) or a vector of the correct length
to replace the diagonal.
lowerOnly Default = FALSE.
See Also
Similar functions exist in many packages, see vec2sm in corpcor, xpnd in MCMCpack

#### Examples
```{r}


x <- 1:6
vech2mat(x)
vech2mat(x, diag = 7)
vech2mat(x, diag = c(99, 98, 97, 96))
vech2mat(x, diag = 0, lowerOnly = TRUE)
waldt 125
