---
title: "``vech2Corr ``"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rockchalk)
```

Convert the vech (column of strictly lower trianglar values from a matrix) into a correlation matrix.

#### Description

vech2Corr is a convenience function for creating correlation matrices from a vector of the lower triangular values. It checks the arguments  to make sure they are consistent with the requirements of a correlation matrix. All values must be in [-1, 1], and the number of values specified must be correct for a lower triangle.

#### Usage
<pre><code>

vech2Corr(vech)

</code></pre>

#### Arguments

* vech A vector of values for the strictly lower triangle of a matrix. All values must be in the [0,1] interval (because they are correlations) and the matrix formed must be positive definite.


#### Examples

```{r}

v <- c(0.1, 0.4, -0.5)
vech2Corr(v)
v <- c(0.1, 0.4, -0.4, 0.4, 0.5, 0.1)
vech2Corr(v)

```

#### Details

Use this in combination with the lazyCov function to convert a vector of standard deviations and the correlation matrix into a covariance matrix.

#### Value
A symmetric correlation matrix, with 1’s on the diagonal.

#### Author(s)
Paul E. Johnson <pauljohn@ku.edu>

#### See Also
Similar functions exist in many packages, see vec2sm in corpcor, xpnd in MCMCpack
