---
title: "``Rockchalk::waldt``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Rockchalk)
```

waldt T-test for the difference in 2 regression parameters

#### Description

This is the one the students call the "fancy t test". It is just the simplest, most easy to use version
of the t test to decide if 2 coefficients are equal. It is not as general as other functions in other
packages. This is simpler to use for beginners. The car package’s function linearHypothesis
is more general, but its documentation is much more difficult to understand. It gives statistically
identical results, albeit phrased as an F test.

#### Usage
<pre><code>


waldt(parm1, parm2, model, model.cov = NULL, digits = getOption("digits"))

</code></pre>
#### Arguments

parm1 A parameter name, in quotes!
parm2 Another parameter name, in quotes!
model A fitted regression model
model.cov Optional, another covariance matrix to use while calculating the test. Primarily
used for robust (or otherwise adjusted) standard errors
digits How many digits to print? This affects only the on-screen printout. The return
object is numeric, full precision.

#### Details

I did this because we have trouble understanding terminology in documentation for more abstract
functions in other R packages.
It has an additional feature, it can import robust standard errors to conduct the test.
Value
A vector with the difference, std. err., t-stat, and p value. Prints a formatted output statement.
Author(s)
Paul Johnson <pauljohn@ku.edu>

#### Examples
```{r}


mdat <- data.frame(x1 = rnorm(100), x2 = rnorm(100))
stde <- 2
mdat$y <- 0.2 * mdat$x1 + 0.24 * mdat$x2 + stde * rnorm(100)
m1 <- lm(y ~ x1 + x2, data = mdat)
waldt("x1", "x2", m1)
waldt("x1", "x2", m1, digits = 2)
126 waldt
## Returned object is not "rounded characters". It is still numbers
stillnumeric <- waldt("x1", "x2", m1, digits = 2)
stillnumeric
## Equivalent to car package linearHypothesis:
if(require(car)){
linearHypothesis(m1, "x1 = x2")
}
## recall t = sqrt(F) for a 1 degree of freedom test.
## If we could understand instructions for car, we probably
## would not need this function, actually.
Index
