ols_aic Akaike in Format ion criterion

#### Description 

Akaike in
#### Format 
ion criterion for model selection.

####Usage 
<pre><code>
  
  ols_aic(model, method = c("R", "STATA", "SAS"))

</code></pre>
  #### Arguments 
  
  model An object of class lm.
method A character vector; specify the method to compute AIC. Valid options include
R, STATA and SAS.

#### Details 

AIC provides a means for model selection. Given a collection of models for the data, AIC estimates
the quality of each model, relative to each of the other models. R and STATA use loglikelihood to
compute AIC. SAS uses residual sum of squares. Below is the formula in each case:
  R & STATA
AIC = 􀀀2(loglikelihood) + 2p
SAS
AIC = n  ln(SSE=n) + 2p
where n is the sample size and p is the number of model parameters including intercept.
Value
Akaike in Formation criterion of the model.
6 ols_apc
References
Akaike, H. (1969). “Fitting Autoregressive Models for Prediction.” Annals of the Institute of Statistical
Mathematics 21:243–247.
Judge, G. G., Griffiths, W. E., Hill, R. C., and Lee, T.-C. (1980). The Theory and Practice of
Econometrics. New York: John Wiley & Sons.
See Also
Other model selection criteria: ols_apc, ols_fpe, ols_hsp, ols_mallows_cp, ols_msep, ols_sbc,
ols_sbic

####Examples 
```{r} 

# using R computation method
model <- lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_aic(model)
```

```{r}
# using STATA computation method
model <- lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_aic(model, method = 'STATA')
```

```{r}
# using SAS computation method
model <- lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_aic(model, method = 'SAS')
```
