---
title: "``ols_sbc``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(olsrr)
```

ols_sbc Bayesian information criterion

#### Description 
 
Bayesian information criterion for model selection.

####Usage 
<pre><code>
ols_sbc(model, method = c("R", "STATA", "SAS"))
</code></pre>

#### Arguments
* ``model``: An object of class lm.
* ``method``: A character vector; specify the method to compute BIC. Valid options include
R, STATA and SAS.


####Examples 
```{r} 

# using R computation method
model <- lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_sbc(model)
# using STATA computation method
model <- lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_sbc(model, method = 'STATA')
# using SAS computation method
model <- lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_sbc(model, method = 'SAS')
```

#### Value
The bayesian information criterion of the model.

#### Details
SBC provides a means for model selection. Given a collection of models for the data, SBC estimates
the quality of each model, relative to each of the other models. R and STATA use loglikelihood to
compute SBC. SAS uses residual sum of squares. Below is the formula in each case:

#### R & STATA
AIC = 􀀀2(loglikelihood) + ln(n)  2p

#### SAS
AIC = n  ln(SSE=n) + p  ln(n)
where n is the sample size and p is the number of model parameters including intercept.



#### References
Schwarz, G. (1978). “Estimating the Dimension of a Model.” Annals of Statistics 6:461–464.
Judge, G. G., Griffiths, W. E., Hill, R. C., and Lee, T.-C. (1980). The Theory and Practice of
Econometrics. New York: John Wiley & Sons.
See Also
Other model selection criteria: ols_aic, ols_apc, ols_fpe, ols_hsp, ols_mallows_cp, ols_msep,
ols_sbic
48 ols_sbic