---
title: "``ols_sbic ``"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(olsrr)
```

### Sawa's bayesian information criterion

#### Description 
 
Sawa's bayesian information criterion for model selection.

####Usage 
<pre><code>

ols_sbic(model, full_model)
</code></pre>

#### Arguments
model An object of class lm.
full_model An object of class lm.


####Examples 
 ```{r} 

full_model <- lm(mpg ~ ., data = mtcars)
model <- lm(mpg ~ disp + hp + wt + qsec, data = mtcars)
ols_sbic(model, full_model)
```

#### Details
Sawa (1978) developed a model selection criterion that was derived from a Bayesian modification of
the AIC criterion. Sawa’s Bayesian Information Criterion (BIC) is a function of the number of observations
n, the SSE, the pure error variance fitting the full model, and the number of independent
variables including the intercept.
SBIC = n  ln(SSE=n) + 2(p + 2)q 􀀀 2(q2)
where q = n(2)=SSE, n is the sample size, p is the number of model parameters including
intercept SSE is the residual sum of squares.
Value
Sawa’s Bayesian Information Criterion


#### References
Sawa, T. (1978). “Information Criteria for Discriminating among Alternative Regression Models.”
Econometrica 46:1273–1282.
Judge, G. G., Griffiths, W. E., Hill, R. C., and Lee, T.-C. (1980). The Theory and Practice of
Econometrics. New York: John Wiley & Sons.
See Also
Other model selection criteria: ols_aic, ols_apc, ols_fpe, ols_hsp, ols_mallows_cp, ols_msep,
ols_sbc

